---
title: "SSPS [4102|6006] Week 11: Uncertainty"
tutorial:
  id: "week-11"
output:
  learnr::tutorial:
    progressive: true
    ace_theme: github
    theme: united
runtime: shiny_prerendered
description: "This tutorial will cover machine learning and text analysis"
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(sspslearnr)
library(gradethis)
library(class)
library(tidytext)

tutorial_options(exercise.checker = gradethis::grade_learnr, 
                 exercise.reveal_solution="TRUE")
knitr::opts_chunk$set(echo = FALSE)
num_tutorial <- 0
num_excercise <- 0
tut_reptitle <- "SSPS [4102|6006] Week 11: In-class Individual Tutorial"
```

## Machine learning

```{r ml-quiz}
quiz(
  question("What is the main goal of supervised learning in machine learning?",
           answer("To discover hidden patterns in data without labels"),
           answer("To predict outcomes based on labeled training data", correct = TRUE),
           answer("To automatically cluster data points"),
           allow_retry = TRUE,
           correct = "Correct! Supervised learning involves predicting outcomes based on labeled data.",
           incorrect = "Not quite. Supervised learning relies on labeled data to make predictions."
  ),
  
  question("Which of the following is a common evaluation metric for classification tasks?",
           answer("Mean Absolute Error"),
           answer("Accuracy", correct = TRUE),
           answer("Root Mean Square Error"),
           allow_retry = TRUE,
           correct = "Correct! Accuracy is a commonly used metric for evaluating classification models.",
           incorrect = "Not quite. For classification tasks, metrics like accuracy or precision are usually applied."
  ),
  
  question("What is overfitting in machine learning?",
           answer("When a model performs well on training data but poorly on new data", correct = TRUE),
           answer("When a model generalizes well to new data"),
           answer("When a model has high bias and low variance"),
           allow_retry = TRUE,
           correct = "Correct! Overfitting occurs when the model performs well on training data but fails to generalize.",
           incorrect = "Not quite. Overfitting refers to a model performing poorly on new data."
  ),
  
  question("Why is it important to separate data into training and test sets?",
         answer("To ensure the model performs well on the training data"),
         answer("To prevent the model from seeing the test data, providing an unbiased evaluation of performance on unseen data", correct = TRUE),
         answer("To reduce the size of the dataset for faster computation"),
         allow_retry = TRUE,
         correct = "Correct! Separating data helps us evaluate model performance on unseen data, reducing bias.",
         incorrect = "Not quite. Separating training and test data ensures that the model's performance is assessed objectively on new data."
)
  
)
```


## R2 and RMSE

```{r r2}
quiz(
  question("What does the R2 metric indicate in a regression model?",
           answer("The percentage of variance in the dependent variable explained by the model", correct = TRUE),
           answer("The average error of predictions"),
           answer("The amount of data points close to the mean"),
           allow_retry = TRUE,
           correct = "Correct! R2 measures how well the model explains the variance in the outcome variable.",
           incorrect = "Not quite. R2 represents the proportion of variance explained by the model."
  ),
  
  question("If a regression model has an R2 of 0, what does this imply?",
           answer("The model perfectly predicts the outcome variable"),
           answer("The model explains none of the variance in the outcome variable", correct = TRUE),
           answer("The model has high accuracy"),
           allow_retry = TRUE,
           correct = "Correct! An R2 of 0 means the model explains none of the variance.",
           incorrect = "Not quite. R2 of 0 implies no variance is explained by the model."
  ),
  
  question("What is the main purpose of the Root Mean Square Error (RMSE) metric?",
           answer("To evaluate the average magnitude of prediction errors", correct = TRUE),
           answer("To indicate how well the model explains the variance in the data"),
           answer("To compare training and test data"),
           allow_retry = TRUE,
           correct = "Correct! RMSE provides the average magnitude of prediction errors in the same units as the target variable.",
           incorrect = "Not quite. RMSE evaluates the size of errors in predictions."
  ),
  
  question("How is RMSE interpreted in the context of model performance?",
           answer("Lower RMSE values indicate a better model fit", correct = TRUE),
           answer("Higher RMSE values indicate better predictions"),
           answer("RMSE is unaffected by the units of the outcome variable"),
           allow_retry = TRUE,
           correct = "Correct! A lower RMSE means the model's predictions are closer to actual values, indicating better performance.",
           incorrect = "Not quite. Lower RMSE values suggest better model fit, as they mean predictions are closer to actual values."
  ),
  
  question("What would a very high RMSE suggest about a model?",
           answer("The model's predictions are far from the actual values", correct = TRUE),
           answer("The model perfectly fits the data"),
           answer("The model explains all the variance in the outcome variable"),
           allow_retry = TRUE,
           correct = "Correct! A high RMSE indicates that the model's predictions deviate significantly from actual values.",
           incorrect = "Not quite. A high RMSE suggests that the model is making large prediction errors."
  )
)
```

## p-values

Let's go back to p-values...

- p-values should be approached with caution and not misinterpreted as the probability of a hypothesis being false. 

- p-value only indicates the likelihood of observing the data given a specific hypothesis, not the likelihood of the hypothesis itself.

```{r p-value}
quiz(
  question("What does a p-value represent in hypothesis testing?",
           answer("The probability of obtaining results as extreme as the observed results, assuming the null hypothesis is true", correct = TRUE),
           answer("The probability that the alternative hypothesis is true"),
           answer("The probability of observing any data"),
           allow_retry = TRUE,
           correct = "Correct! The p-value measures the probability of observing data as extreme as the observed data if the null hypothesis is true.",
           incorrect = "Not quite. The p-value represents the likelihood of observing data as extreme as the observed results under the null hypothesis."
  ),
  
  question("If a p-value is 0.03, what does this imply if we have a significance level of 0.05?",
           answer("There is strong evidence to reject the null hypothesis", correct = TRUE),
           answer("The null hypothesis should be accepted"),
           answer("There is no evidence against the null hypothesis"),
           allow_retry = TRUE,
           correct = "Correct! A p-value of 0.03 is less than 0.05, suggesting strong evidence against the null hypothesis.",
           incorrect = "Not quite. A p-value of 0.03 is below the 0.05 significance level, indicating evidence to reject the null hypothesis."
  ),
  
  question("What does a large p-value (e.g., 0.5) indicate about the null hypothesis?",
           answer("It suggests weak evidence against the null hypothesis, so it should not be rejected", correct = TRUE),
           answer("It suggests strong evidence against the null hypothesis"),
           answer("It suggests that the null hypothesis is definitely true"),
           allow_retry = TRUE,
           correct = "Correct! A high p-value indicates weak evidence against the null hypothesis, suggesting it should not be rejected.",
           incorrect = "Not quite. A large p-value implies there is not enough evidence to reject the null hypothesis."
  ),
  
  question("What is a common threshold (alpha) for determining statistical significance with p-values?",
           answer("0.01"),
           answer("0.05", correct = TRUE),
           answer("0.5"),
           allow_retry = TRUE,
           correct = "Correct! 0.05 is a commonly used threshold for statistical significance.",
           incorrect = "Not quite. While thresholds can vary, 0.05 is often used as the significance level in hypothesis testing."
  ),
  
  question("Does a small p-value (e.g., < 0.01) prove that the alternative hypothesis is true?",
           answer("Yes, it proves the alternative hypothesis is true"),
           answer("No, it only suggests that the data is unlikely under the null hypothesis", correct = TRUE),
           answer("Yes, it guarantees that the null hypothesis is false"),
           allow_retry = TRUE,
           correct = "Correct! A small p-value does not prove the alternative hypothesis; it only suggests that the null hypothesis is unlikely.",
           incorrect = "Not quite. A small p-value means the data is unlikely under the null hypothesis but does not prove the alternative hypothesis."
  )
)
```

## Reporting on hypothesis testing

```{r htest}
quiz(
  question("How should you start a conclusion statement when reporting the results of hypothesis testing?",
           answer("By explaining the context of the hypothesis and the assumptions made", correct = TRUE),
           answer("By stating the test statistic value"),
           answer("By only mentioning the sample data"),
           allow_retry = TRUE,
           correct = "Correct! Begin by explaining the context of the hypothesis and any assumptions.",
           incorrect = "Not quite. The conclusion should begin with the context and assumptions."
  ),
  
  question("What should you include in the conclusion about the estimated effect?",
           answer("Only whether the effect is positive or negative"),
           answer("Both the direction, size of the effect, and unit of measurement", correct = TRUE),
           answer("Only the statistical significance of the effect"),
           allow_retry = TRUE,
           correct = "Correct! Report the effect's direction, size, and unit to clearly communicate the result.",
           incorrect = "Not quite. Mention the effect's direction, size, and unit for clarity."
  ),
  
  question("If the p-value is below 0.05, what should you report about the null hypothesis?",
           answer("We fail to reject the null hypothesis"),
           answer("We reject the null hypothesis, indicating the effect is statistically significant", correct = TRUE),
           answer("The null hypothesis is accepted"),
           allow_retry = TRUE,
           correct = "Correct! A p-value below 0.05 means we reject the null hypothesis and report statistical significance.",
           incorrect = "Not quite. A low p-value means we reject the null hypothesis and report significance."
  ),
  
  question("How should statistical significance be interpreted when reporting results?",
           answer("If significant, it likely indicates a non-zero effect at the population level", correct = TRUE),
           answer("Significance proves the true effect is exactly equal to the estimated effect"),
           answer("It indicates the model has no error"),
           allow_retry = TRUE,
           correct = "Correct! Statistical significance suggests the effect is likely non-zero at the population level.",
           incorrect = "Not quite. Statistical significance means the effect is unlikely to be zero, not that the estimate is exact."
  ),
  
  question("What is the importance of replication in hypothesis testing results?",
           answer("It increases confidence by showing similar conclusions across samples", correct = TRUE),
           answer("It allows the null hypothesis to be accepted with certainty"),
           answer("It’s only necessary if the p-value is greater than 0.05"),
           allow_retry = TRUE,
           correct = "Correct! Replication confirms results by showing similar findings in different samples.",
           incorrect = "Not quite. Replication helps confirm findings but doesn't 'prove' any hypothesis."
  )
)
```

## Regression analysis in R

```{r reg}

quiz(
  question("What function is commonly used in R to fit a linear regression model?",
           answer("lm()", correct = TRUE),
           answer("glm()"),
           answer("fit()"),
           allow_retry = TRUE,
           correct = "Correct! The lm() function is used for fitting linear regression models.",
           incorrect = "Not quite. The lm() function is specifically used for linear regression."
  ),
  
  question("What argument should you set in the glm() function to specify a logistic regression model?",
           answer("family = 'logistic'"),
           answer("family = binomial", correct = TRUE),
           answer("type = 'binary'"),
           allow_retry = TRUE,
           correct = "Correct! Setting family = binomial in glm() specifies a logistic regression model.",
           incorrect = "Not quite. Use family = binomial for logistic regression in glm()."
  ),
  
  question("Which evaluation metric is commonly used to assess logistic regression model performance?",
           answer("Root Mean Square Error (RMSE)"),
           answer("Accuracy", correct = TRUE),
           answer("Adjusted R-squared"),
           allow_retry = TRUE,
           correct = "Correct! Accuracy is a common metric for evaluating logistic regression models.",
           incorrect = "Not quite. Accuracy is typically used for logistic regression evaluation."
  ),
  
  question("What is the purpose of using the summary() function on a linear or logistic regression model?",
           answer("To fit the model to the data"),
           answer("To view detailed results and statistics from the model fit", correct = TRUE),
           answer("To create predictions for new data"),
           allow_retry = TRUE,
           correct = "Correct! The summary() function provides detailed statistics on the fitted model.",
           incorrect = "Not quite. summary() is used to view statistics on the fitted model."
  ),
  
  question("In a logistic regression model, what does an odds ratio greater than 1 indicate?",
           answer("The predictor is associated with a lower likelihood of the outcome"),
           answer("The predictor is associated with a higher likelihood of the outcome", correct = TRUE),
           answer("The predictor has no impact on the outcome"),
           allow_retry = TRUE,
           correct = "Correct! An odds ratio greater than 1 suggests a higher likelihood of the outcome.",
           incorrect = "Not quite. An odds ratio above 1 indicates an increased likelihood."
  ),
  
  question("What R function is used to make predictions from a regression model on new data?",
           answer("predict()", correct = TRUE),
           answer("summary()"),
           answer("estimate()"),
           allow_retry = TRUE,
           correct = "Correct! The predict() function is used to generate predictions for new data.",
           incorrect = "Not quite. Use predict() for making predictions with a fitted model."
  ),
  
  question("What does the coefficient of a predictor variable represent in a linear regression model?",
           answer("The expected change in the outcome variable for a one-unit increase in the predictor", correct = TRUE),
           answer("The probability of the outcome given the predictor"),
           answer("The proportion of variance explained by the model"),
           allow_retry = TRUE,
           correct = "Correct! In linear regression, the coefficient represents the change in the outcome for a one-unit change in the predictor.",
           incorrect = "Not quite. In linear regression, the coefficient indicates the expected change in the outcome."
  )
)
```

### Now using the `civilwar` data set and R calculate the effect of `polity2` (democracy) and `ef`(ethnic frammentation) on `war` (civil war is ongoing that year if `1`). What is the most appropriate model to measure the effect on `war`?

```{r estimate, exercise = TRUE, exercise.blanks = "___+"}
# Load the data (don't change this)
civilwar <- read.csv("data/civilwar.csv")

# Now on the line below fit your model predicting `war`
fit <- ___(war ~ ___ data = civilwar)

# Finally, display the summary of the model
summary(fit)
```

```{r estimate-solution}
civilwar <- read.csv("data/civilwar.csv")

fit <- glm(war ~ polity2 + ef, family = "binomial" , data = civilwar)

summary(fit)
```

```{r estimate-check}
grade_code()
```

## Submit Report

```{r context="setup"}
submission_ui
```

```{r context="server"}
submission_server()
```
