---
title: "SSPS [4102|6006] Week 09: Uncertainty"
tutorial:
  id: "week-09"
output:
  learnr::tutorial:
    progressive: true
    ace_theme: github
    theme: united
runtime: shiny_prerendered
description: "This tutorial will cover `knn()`, `lm()` and `glm()`"
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(sspslearnr)
library(gradethis)
library(class)

tutorial_options(exercise.checker = gradethis::grade_learnr, 
                 exercise.reveal_solution="TRUE")
knitr::opts_chunk$set(echo = FALSE)
num_tutorial <- 0
num_excercise <- 0
tut_reptitle <- "SSPS [4102|6006] Week 09: In-class Individual Tutorial"
```

## Quiz: Supervised vs. Unsupervised Learning

```{r supervised_unsupervised_quiz, echo=FALSE}
quiz(
  question("What is a key characteristic of supervised learning?",
           answer("It does not require any data"),
           answer("It requires labeled data for training", correct = TRUE),
           answer("It only works with numerical data"),
           allow_retry = TRUE,
           correct = "Correct! Supervised learning requires labeled data to train the model.",
           incorrect = "Not quite. Supervised learning specifically needs labeled data for training."),
  
  question("Which of the following is an example of unsupervised learning?",
           answer("Clustering customers based on purchasing behavior", correct = TRUE),
           answer("Predicting house prices based on features"),
           answer("Classifying emails as spam or not spam"),
           allow_retry = TRUE,
           correct = "Correct! Clustering customers is a classic example of unsupervised learning.",
           incorrect = "Not quite. Clustering customers is an example of unsupervised learning, while the others are supervised.")
)
```

## Quiz: Parametric vs. Non-Parametric Learning

```{r parametric_nonparametric_quiz, echo=FALSE}
quiz(
  
    question("What is a defining feature of parametric learning methods?",
           answer("They use a fixed number of coefficients to model the relationship", correct = TRUE),
           answer("They adapt the model based on the data without fixed parameters"),
           answer("They do not rely on any mathematical functions"),
           allow_retry = TRUE,
           correct = "Correct! Parametric methods use a fixed set of coefficients to define the model.",
           incorrect = "Not quite. Parametric methods involve a predetermined number of coefficients to describe the relationship."),

  question("Which method does not rely on fixed coefficients to define its model?",
           answer("Parametric learning"),
           answer("Non-parametric learning", correct = TRUE),
           answer("Both parametric and non-parametric learning"),
           allow_retry = TRUE,
           correct = "Correct! Non-parametric methods do not use fixed coefficients; they adapt based on the data.",
           incorrect = "Not quite. Non-parametric methods do not assume a fixed set of coefficients and adapt to the data directly."),

  question("In parametric models, what do the coefficients represent?",
           answer("The relationship between predictors and the outcome variable", correct = TRUE),
           answer("The distance between data points"),
           answer("The number of data points used in training"),
           allow_retry = TRUE,
           correct = "Correct! In parametric models, coefficients quantify the relationship between predictors and the outcome.",
           incorrect = "Not quite. Coefficients in parametric models reflect the relationship between predictors and the outcome."),

  question("Why might non-parametric methods require more data than parametric methods?",
           answer("They are always more complex and cannot handle simple relationships"),
           answer("They do not rely on fixed coefficients, needing more data to capture relationships", correct = TRUE),
           answer("They only work with very large datasets"),
           allow_retry = TRUE,
           correct = "Correct! Non-parametric methods adapt their models based on the data, often needing more samples to accurately capture relationships.",
           incorrect = "Not quite. Non-parametric methods need more data to effectively model the relationships without fixed coefficients."),
  
  question("Which of the following is an example of a non-parametric method?",
           answer("Linear Regression"),
           answer("k-Nearest Neighbors (k-NN)", correct = TRUE),
           answer("Logistic Regression"),
           allow_retry = TRUE,
           correct = "Correct! k-NN is a classic example of a non-parametric learning method.",
           incorrect = "Not quite. k-NN is non-parametric, while linear and logistic regression are parametric methods.")
)
```



## Quiz: Key Concepts of k-NN

```{r knn-quiz}
quiz(
  question("What does the 'k' in k-NN stand for?",
           answer("The number of features in the dataset"),
           answer("The number of nearest neighbors", correct = TRUE),
           answer("The number of classes"),
           answer("The number of samples"),
           allow_retry = TRUE,
           correct = "Correct! 'k' refers to the number of nearest neighbors used in the classification.",
           incorrect = "Not quite. The 'k' in k-NN refers to the number of nearest neighbors used."),
  
  question("What type of learning method is k-NN?",
           answer("Eager learning"),
           answer("Supervised learning"),
           answer("Lazy learning", correct = TRUE),
           answer("Unsupervised learning"),
           allow_retry = TRUE,
           correct = "Correct! k-NN is a lazy learning algorithm because it doesn't build an explicit model.",
           incorrect = "Not quite. k-NN is a lazy learning method because it stores training data without building an explicit model."),
  
  question("Which distance measure is most commonly used in k-NN?",
           answer("Manhattan distance"),
           answer("Cosine similarity"),
           answer("Euclidean distance", correct = TRUE),
           answer("Jaccard distance"),
           allow_retry = TRUE,
           correct = "Correct! Euclidean distance is commonly used to measure similarity in k-NN.",
           incorrect = "Not quite. The most common distance measure in k-NN is Euclidean distance.")
)
```

## Predict with `knn()`

Let's consider this dataset of countries with two 2020 country-level measures collected by the World Bank.

```{r knn-load}
countries <- 
  read.csv("data/countries.csv") |>
  dplyr::filter(country != "Mauritius")
head(countries)
```

- `NY.GDP.PCAP.KD`: GDP per capita (constant 2015 US$)
- `SH.XPD.CHEX.GD.ZS`: Current health expenditure (% of GDP)

The dataset `countries` doesn't contain the record for Mauritius, which is this: 

```{r}
head(read.csv("data/countries.csv") |>
  dplyr::filter(country == "Mauritius"))
```

### Use knn() to train your classifier on the dataframe `countries` and predict the `region` for Mauritius.

- Note you will need to set the values for the predictors for the country from above.

```{r knn-ex-1, exercise = TRUE, exercise.blanks = "___+", exercise.setup = "knn-load"}
knn(train = dplyr::select(countries, NY.GDP.PCAP.KD, SH.XPD.CHEX.GD.ZS),
    test = data.frame(___),
    cl = ___,
    k = 3 # Let's set this to 3
) 
```

```{r knn-ex-1-solution}
knn(train = dplyr::select(countries, NY.GDP.PCAP.KD, SH.XPD.CHEX.GD.ZS),
    test = data.frame(NY.GDP.PCAP.KD = 9362.874, SH.XPD.CHEX.GD.ZS = 6.482102),
    cl = countries$region,
    k = 3 # Let's set this to 3
)
```

```{r knn-ex-quiz}
quiz(
  question("What region does knn predict for Mauritius?",
           answer("East Asia & Pacific"),
           answer("Europe & Central Asia"),
           answer("Latin America & Caribbean", correct = TRUE),
           answer("Middle East & North Africa"),
           answer("North America"),
           answer("South Asia"),
           answer("Sub-Saharan Africa"),
           allow_retry = TRUE))
```


## Quiz: Key Concepts of Linear Regression

```{r linear_regression_quiz, echo=FALSE}
quiz(
  question("What does the coefficient in a linear regression represent?",
           answer("The standard error of the predictor"),
           answer("The change in the predictor for a one-unit change in the outcome"),
           answer("The change in the outcome variable for a one-unit change in the predictor", correct = TRUE),
           allow_retry = TRUE,
           correct = "Correct! The coefficient represents the change in the outcome for each one-unit change in the predictor.",
           incorrect = "Not quite. The coefficient shows how the outcome changes with the predictor."),
  
  question("What is the purpose of the residuals in a linear regression model?",
           answer("To determine the significance of the predictors"),
                      answer("To measure the difference between observed and predicted values", correct = TRUE),
           answer("To improve the model's accuracy"),
           allow_retry = TRUE,
           correct = "Correct! Residuals represent the difference between the observed and predicted values.",
           incorrect = "Not quite. Residuals measure the error between observed and predicted values.")
)
```

## Quiz: Key Concepts of Logistic Regression

```{r logistic_regression_quiz, echo=FALSE}
quiz(
  question("What type of outcome does logistic regression predict?",
           answer("Continuous outcomes"),
           answer("Binary outcomes (0/1)", correct = TRUE),
           answer("Categorical outcomes with more than two categories"),
           allow_retry = TRUE,
           correct = "Correct! Logistic regression is used for predicting binary outcomes.",
           incorrect = "Not quite. Logistic regression is specifically for binary outcomes."),
  
  question("What does the logistic regression coefficient represent?",
           answer("The change in the probability of the outcome for a one-unit change in the predictor"),
           answer("The average effect of the predictor on the outcome"),
           answer("The change in log-odds of the outcome for a one-unit change in the predictor", correct = TRUE),
           allow_retry = TRUE,
           correct = "Correct! The coefficient represents the change in log-odds with a one-unit predictor change.",
           incorrect = "Not quite. Logistic regression coefficients reflect changes in log-odds.")
)
```

## Logistic regression with `glm()`

Consider this data set

```{r load-logit}
adminissions <- 
  read.csv("data/binary.csv")
head(adminissions)
```

This dataset `adminissions` includes a binary response (outcome, dependent) variable called `admit`. There are three predictor variables: `gre`, `gpa`, and `rank`. The variables `gre` and `gpa` will be treated as continuous. The variable `rank` ranges from 1 to 4, where institutions with a `rank` of 1 have the highest prestige, and those with a `rank` of 4 have the lowest. Note that `rank` is loaded as a numeric variable, you must consider it as a factor in your logistic regression formula (`factor(rank)`).

```{r logit-ex, exercise = TRUE, exercise.setup = "load-logit", exercise.blanks = "___+"}
fit <- glm(formula = ___, data = ___, family = ___)
summary(fit)
```

```{r logit-ex-solution}
fit <- glm(formula = admit ~ gre + gpa + factor(rank), data = binary, family = binomial)
summary(fit)
```

```{r logit-ex-1, echo = FALSE}
quiz(
  question("Is the `gre` result positively or negatively associated with the probability of admission?",
           answer("Positively", correct = TRUE),
           answer("Negatively"),
           allow_retry = TRUE,
           correct = "Correct! The estimated log-odds is +0.002264.",
           incorrect = "Not quite. The estimated log-odds is +0.002264."),
  question("Is the `gre` result significantly associated with the probability of admission at the 5% level?",
           answer("Yes", correct = TRUE),
           answer("No"),
           allow_retry = TRUE,
           correct = "Correct! The p-value for `gre` is 0.038465, which is smaller than 0.05.",
           incorrect = "Not quite. The p-value for `gre` is 0.038465, which is smaller than 0.05."),
  question("Is an increase of 1 point in a student's `gpa` associated on average with an increase of 80 percentage points in their chance of adminission?",
           answer("Yes"),
           answer("No", correct = TRUE),
           allow_retry = TRUE,
           correct = "Correct! The estimates of a logistic regression are not expressed in terms of the probability of the outcome but in log-odds.",
           incorrect = "Not quite. The estimates of a logistic regression are not expressed in terms of the probability of the outcome but in log-odds."))
```


## Submit Report

```{r context="setup"}
submission_ui
```

```{r context="server"}
submission_server()
```
