---
title: "SSPS [4102|6006] Week 04: Data visualisation"
tutorial:
  id: "week-04"
output:
  learnr::tutorial:
    progressive: true
    ace_theme: github
    theme: united
runtime: shiny_prerendered
description: "This tutorial will cover basic descriptive statistics and
  data visualization. (Tutorial designed by Matt Blackwell)"
---

```{r setup, include=FALSE}
library(gradethis)
library(learnr)
library(sspslearnr)
library(ggplot2)

tutorial_options(exercise.checker = gradethis::grade_learnr, 
                 exercise.reveal_solution="TRUE")
knitr::opts_chunk$set(echo = FALSE)
num_tutorial <- 0
num_excercise <- 0
tut_reptitle <- "SSPS [4102|6006] Week 04: In-class Individual Tutorial"

# Data
data(STAR, package = "qss")
star <- STAR

```

## Handling Missing Data in R

### Small class size data


In this tutorial, you'll analyse data from the STAR project, which is a four-year randomized trial on the effectiveness of small class sizes on education performance. The `star` data frame as been loaded into your space so that you can play around with it a bit.

#### Exercises
- Use the `head` function on the `star` to see what the data looks like.

```{r head-star, exercise = TRUE}

```

```{r head-star-solution}
head(star)
```

```{r head-star-check}
grade_code()
```

- Use the `dim` function on the `star` to see what the dimensions of the data look like.

```{r dim-star, exercise = TRUE}

```

```{r dim-star-solution}
dim(star)
```

```{r dim-star-check}
grade_code()
```


- Use the `summary` function on the `star` to get a sense for each variable.

```{r summary-star, exercise = TRUE}

```

```{r summary-star-solution}
summary(star)
```

```{r summary-star-check}
grade_code()
```



### Handling missing data


You probably noticed that there were some `NA` values in the data when you used the `head()` function. These are **missing values**, where the value for that unit on that variable is missing or unknown. These values pose problems when we are trying to calculate quantities of interest like means or medians because R doesn't know how to handle them.

The first tool in your toolkit for missing data is the `is.na()` function. When you pass a vector `x` to `is.na(x)`, it will return a vector of the same length where each entry is `TRUE` if the value of `x` is `NA` and `FALSE` otherwise. Using logicals, you can easily get the opposite vector `!is.na(x)` which is `TRUE` when `x` is observed and `FALSE` when `x` is missing.

#### Exercises
- Use the `is.na` and `head` functions to show whether or not the first 6 values of the `g4math` variable from the `star` data frame are missing.

```{r headna, exercise = TRUE}

```

```{r headna-solution}
head(is.na(star$g4math))
```

```{r headna-check}
grade_result(
  pass_if(~ identical(.result, head(is.na(star$g4math))))
)
```

Remember that you can sum a vector of logicals with `TRUE` being counted as a `1` and `FALSE` being counted as a `0`. Try to run this

```{r check-sum-logicals, exercise = TRUE}
sum(c(TRUE, FALSE, TRUE, FALSE))
```

- Use the `is.na` and `sum` functions to show how many values of the `g4math` variable are missing.


```{r sumna, exercise = TRUE}

```

```{r sumna-solution}
sum(is.na(star$g4math))
```

```{r sumna-check}
grade_code()
```


- Use the `is.na` and `mean` functions to show what proportion of the `g4math` variable is missing.


```{r meanna, exercise = TRUE}

```

```{r meanna-solution}
mean(is.na(star$g4math))
```

```{r meanna-check}
grade_code()
```



### Calculating means in the fact of missing data

Missing values makes it difficult to calculate numerical quantities of interest like the mean, median, standard deviation, or variance. Many of these function will simply return `NA` if there is a single missing value in the vector. We can instruct many function to ignore the missing values and do their calculation on just the observed data by using the `na.rm = TRUE` argument. For instance, suppose we have `x <- c(NA, 1,2,3)`, then `mean(x)` will return `NA`, but `mean(x, na.rm = TRUE)` will return `2`.

#### Exercises
- Try to calculate the `mean` of the `g4math` variable in the `star` data frame without setting `na.rm`.

```{r mean-with-na, exercise = TRUE}

```

```{r mean-with-na-solution}
mean(star$g4math)
```

```{r mean-with-na-check}
grade_code(
  correct = "This isn't that useful though!"
)
```

- Try to calculate the `mean` of the `g4math` variable when setting `na.rm = TRUE` within the function `mean()`.

```{r mean-narm, exercise = TRUE}

```

```{r mean-narm-solution}
mean(star$g4math, na.rm = TRUE)
```

```{r mean-narm-check}
grade_code()
```

## Visualising Data

### Barplots

The **barplot** is a useful way to visualize a categorical or factor variable. In this exercise, you are going to visualiae the `classtype` variable from the `star` data frame, which can take on the following values:

- `1` = small class
- `2` = regular class
- `3` = regular class with aid

#### Exercises
- Use the `ggplot()` and `geom_bar()` functions to plot the frequency of each category of the `classtype` in the `star` data.

Note: With ggplot2 data and aesthetic mappings are passed from the first layer to the last. Although, you can define data and aesthetics on different layers in order for your code to be automatically checked as "correct" define data and aesthetics in the layer zero so in `ggplot(data = ..., aes(...))` and not in `geom_xxx()`. 

```{r barplot, exercise = TRUE}

```

```{r barplot-solution}
ggplot(star, aes(x = classtype)) + geom_bar()
```

```{r barplot-check}
grade_code()
```

### Histograms

For quantitative (numerical) variables, the `geom_bar()` won't work because there are too many unique values. In this case, you will often use a histogram to visualize the a numerical variable.

#### Exercises
- Use the `geom_histogram()` function to create a histogram for the `g4math` variable in the `star` data frame.
- Remember that you can access a particular variable using the `$`.

```{r hist, exercise = TRUE}
## create a histogram of g4math

```


```{r hist-solution}
ggplot(data = star, aes(x = g4math)) + geom_histogram()
```

```{r hist-check}
grade_code("Great job, though the graph is a bit spartan. Let's make it more readable.")
```

You should have received this warning message:

> Removed 3930 rows containing non-finite outside the scale range (`stat_bin()`). 

```{r quiz-missing-data-warning}
quiz(
  question("Why did you get this message?",
    answer("Because there are not enough bins"),
    answer("Because of the limits set by scale_x_continuous()"),
    answer("Because there are missing values in my data", correct = TRUE),
    incorrect = "Incorret: ggplot2 is informing you that 3930 rows from your data frame coudn't be visualised because they contained missing data",
    allow_retry = TRUE
  ))
```


### Sprucing up the histogram

There are several layers you can pass after to `geom_histogram()` that will improve its readability:

- with `title = "your title"` and `x = "your x-axis label"` within the function `labs()` you can include the title and change the label for the x-axis. 
- with `c(low_limit,upper_limit)` within the function `scale_x_continuous()` or `scale_y_continuous()` you can define the limits for the x-axis and y-axis respectively.

(Remember to add all additional layers with a `+`, also set the layers following the order of the instructions and specify data and aes in the layer zero `ggplot(data = my_data, aes(...)`)

#### Exercises
- Create a histogram (with `geom_histogram()`) then set
    - "Distribution of fourth-grade math scores" as title;
    - "Score" as label of the x-asis;
    - Set the limit of the x-axis to 600 and 800.

```{r hist-nice, exercise = TRUE}
## create the histogram with the specifications given in the instructions

```


```{r hist-nice-solution}
ggplot(data = star, aes(x = g4math)) + geom_histogram() + labs(title = "Distribution of fourth-grade math scores", x = "Score") + scale_x_continuous(limits = c(600,800))
```

### Adding lines to a plot

We'll often want to add more information to a plot to make it even more readable. You can do that with commands that add to the current plot, such as `geom_vline()`.
`geom_vline(xintercept = 1)` will add a vertical line to the plot at the specified value (`1` in this example).

#### Exercises
- Use the `geom_vline()` function to add a vertical line at the mean of the `g4math` variable from the `star` data. Remember, there are missing values in that variable so be sure to use the `na.rm = TRUE` argument to drop them.

```{r hist-line, exercise = TRUE}
ggplot(data = star, aes(x = g4math)) + geom_histogram() + labs(title = "Distribution of fourth-grade math scores", x = "Score") + scale_x_continuous(limits = c(600,800)) 

## add a vertical line at the mean of the variable

```

```{r hist-line-solution}
ggplot(data = star, aes(x = g4math)) + geom_histogram() + labs(title = "Distribution of fourth-grade math scores", x = "Score") + scale_x_continuous(limits = c(600,800)) + geom_vline(xintercept = mean(star$g4math, na.rm = T))
```

```{r hist-line-check}
grade_code()
```

## Conceptual questions Part 1

```{r conceptual-quiz-1}
quiz(
       question("To estimate the correlation between two variables x and y, we need to know the mean of each as well as the:",
              answer("maximum"),
              answer("standard deviation", correct = TRUE),
              answer("length")),
     question("What type of relationships does correlation measure?",
              answer("linear relationships", correct = TRUE),
              answer("nonlinear relationships"),
              answer("both")),
  question("What is the main reason researchers use samples instead of collecting data from an entire population?",
    answer("It is easier to analyze smaller datasets"),
    answer("Collecting data from every individual in a population is often too expensive or infeasible", correct = TRUE),
    answer("Samples provide more accurate data than full population surveys"),
    answer("It is the standard practice without any specific reason"),
    incorrect = "The correct answer is that it is often too expensive or infeasible to collect data from the entire population.",
    allow_retry = TRUE
  ),

  question("Why is it important for a sample to be representative of the population?",
    answer("It increases the sample size"),
    answer("It reduces the cost of data collection"),
    answer("It ensures that inferences made about the population from the sample are accurate", correct = TRUE),
    answer("It speeds up the data analysis process"),
    incorrect = "A representative sample ensures that the inferences made about the population are accurate.",
    allow_retry = TRUE
  ),

  question("Which of the following best describes random sampling?",
    answer("Randomly assigning treatments to different groups"),
    answer("Randomly selecting individuals from a population to ensure representativeness", correct = TRUE),
    answer("Randomly shuffling data before analysis"),
    answer("Randomly assigning variables to a dataset"),
    incorrect = "Random sampling involves selecting individuals from a population at random.",
    allow_retry = TRUE
  ),

  question("What is a histogram used for in data analysis?",
    answer("To show the relationship between two variables"),
    answer("To display the frequency distribution of a single variable", correct = TRUE),
    answer("To compute the mean of a dataset"),
    answer("To perform random sampling"),
    incorrect = "A histogram displays the frequency distribution of a single variable.",
    allow_retry = TRUE
  ),

  question("Which of the following statements is true?",
    answer("Random treatment assignment ensures that the sample is representative of the population."),
    answer("Random sampling ensures that the treatment and control groups are comparable."),
    answer("Random treatment assignment is used to estimate causal effects, while random sampling is used to infer population characteristics.", correct = TRUE),
    answer("Random treatment assignment and random sampling are the same concepts."),
    incorrect = "The correct statement is that random treatment assignment is used to estimate causal effects, while random sampling is used to infer population characteristics.",
    allow_retry = TRUE
  ),

  question("What should the sum of the proportions in a table of proportions equal to?",
    answer("0"),
    answer("0.5"),
    answer("1", correct = TRUE),
    answer("It depends on the data"),
    incorrect = "The sum of the proportions in a table of proportions should equal 1.",
    allow_retry = TRUE
  ),

  question("What does a larger standard deviation indicate about a dataset?",
    answer("The data points are closely clustered around the mean"),
    answer("The data points are spread out farther from the mean", correct = TRUE),
    answer("The data has fewer outliers"),
    answer("The data has a perfect normal distribution"),
    incorrect = "A larger standard deviation indicates that the data points are spread out farther from the mean.",
    allow_retry = TRUE
  )

)
```

## More visualization


We'll continue to analyze data from the STAR project, which is a four-year randomized trial on the effectiveness of small class sizes on education performance. The `star` data frame as been loaded into your space so that you can play around with it a bit.

### Boxplots

Boxplots are useful tools to visualize how the distribution of a continuous variable changes across levels of a categorical variable.

With ggplot2 we create a boxplot with `geom_boxplot()`. Importantly `geom_boxplot()` requires two aesthetics (`x` and `y`) where one must be a categorical variable and the other a continuous variable. 

Let's try this now:

```{r boxplot-1, exercise = TRUE}
ggplot(data = star, aes(x = classtype, y = g4math)) + geom_boxplot()
```

What's going on? Why is there only on `classtype` on the x-axis? The problem is that `classtype` is coded in the data frame as a numerical variable. You can check it with

```{r boxplot-2, exercise = TRUE}
class(star$classtype)
```

and

```{r boxplot-3, exercise = TRUE}
head(star$classtype)
```

But there is a solution. You can force R to consider your numerical variable as a character or better as a factor with `factor()`.

Try this

```{r boxplot-4, exercise = TRUE}
factor(star$classtype)
```

`factors()` allows you to define the labels of your levels in this case (1, 2, and 3). Since we now that 1 = "Small class", 2 = "Regular class", 3 = "Regular class with aid", we can do

Try this

```{r boxplot-5, exercise = TRUE}
factor(star$classtype, 
          labels = c("Small class", "Regular class", "Regular class with aid"))
```

#### Exercises

- Produce a boxplot with `geom_boxplot()` where on the x-axis you have the levels of `classtype` and on the y-axis you have the distribution of `g4math`.
- Make sure you specify the labels of your levels.
- Include some informative label for both the x-axis and y-axis: "Class type" and "Score".

Note: when you add labels to your factor's levels make sure that they are in the right order!

```{r boxplot-6, exercise = TRUE}
ggplot(data = star, aes(x = factor(..., labels = c(...)),  y = ...)) + 
  geom_boxplot() +
  labs(...)
```


```{r boxplot-6-solution}
ggplot(data = star, aes(x = factor(classtype, 
                                      labels =  c("Small class", "Regular class", "Regular class with aid")),
                        y = g4math)) + 
  geom_boxplot() +
  labs(x = "Class type", y = "Score")
```

```{r boxplot-6-check}
grade_code()
```

#### Bonus on factor's labels

Note that you can also obtain the same result by setting the label at in the `scale` layer in ggplot, like this

```{r eval = F}
# Don't run
scale_x_discrete(labels = c(
  "1" = "Small class", 
  "2" = "Regular class", 
  "3" = "Regular class with aid"))
```


### Scatter plots

Now you'll think more about to measure bivariate relationships---that is, the relationship between two variables. The `ggplot` function takes the `data` argument followed by `aes` to specify the variable-of-interests, more importantly, specify `geom_point` to create a scatter plot as such: `ggplot(data, aes(x = variable, y = variable)) + geom_point()`. You will use this plot to explore the relationship between math and reading test scores in the `star` data.

#### Exercises
- Create a scatter plot using the `ggplot` and `geom_point` function with `g4math` on the x-axis and `g4reading` on the y-axis, both from the `star` data.

```{r scatter, exercise = TRUE}
## produce a scatterplot of g4math on the x-axis and g4reading on the y-axis

```

```{r scatter-solution}
## produce a scatterplot of g4math on the x-axis and g4reading on the y-axis
ggplot(star, aes(x = g4math, y = g4reading)) +
  geom_point()
```

```{r scatter-check}
grade_code("OK, great plot! Let's make it a bit more polished.")
```


### Plotting two sets of points

Often we want to plot certain points differently than others. For instance, maybe you want to see how the relationship between math and reading scores differs between students in small classes versus those not in small classes. 

#### Exercises

- With `geom_point()`, plot the relationship between `g4math` (on the x-axis) and `g4reading` on the y-axis.
- This time, differentiate data points on the variable `classtype` by specifying `colour` in the `aes()` nested in `geom_point` (Note you might want to consider `classtype` as a factor variable). 


```{r plot-points, exercise = TRUE}

```


```{r plot-points-solution}
ggplot(star, aes(x = g4math, y = g4reading, colour = factor(classtype))) + 
  geom_point() 
  
```

```{r plot-points-check}
grade_this_code()
```

### Finalizing your scatter plot

The scatter plot is looking very good, but it could use a little bit of polish. Let's add axis labels and a title. 

#### Exercises

- To do so, take the current plot and pass it to the `labs` function to make the graph more informative.

```{r niceplot, exercise = TRUE}

```

```{r niceplot-solution}
ggplot(data = star, aes(x = g4math, y = g4reading)) +
  geom_point(aes(color = classtype)) +
  labs(x = "Fourth Grade Math Scores", 
       y = "Fourth Grade Reading Score", 
       title = "Math vs Reading")
```


## Correlation

- How would you calculatge correlation between `g4math` and `g4reading` from the `star` data frame?

Note: You must use the the attribute `use = "complete.obs"` as you have missing values in your variables.

```{r cor-1, exercise = TRUE}
...(... ,...,  use = "complete.obs")
```

```{r cor-1-solution}
cor(star$g4math, star$g4reading, use = "complete.obs")
```

```{r cor-1-check}
grade_code()
```

## Conceptual questions Part 2

```{r conceptual-quiz-2}
quiz(
  question("What does a correlation coefficient of -0.8 indicate?",
    answer("A strong negative linear relationship between two variables", correct = TRUE),
    answer("A strong positive linear relationship between two variables"),
    answer("No linear relationship between two variables"),
    answer("A weak negative linear relationship between two variables"),
    incorrect = "A correlation coefficient of -0.8 indicates a strong negative linear relationship.",
    allow_retry = TRUE
  ),
    question("Which of the following best describes the purpose of a scatter plot?",
    answer("To show the distribution of a single variable"),
    answer("To compare the mean and median of a dataset"),
    answer("To visualize the relationship between two variables", correct = TRUE),
    answer("To calculate the correlation coefficient"),
    incorrect = "A scatter plot is used to visualize the relationship between two variables.",
    allow_retry = TRUE
  ),
    question("In a skewed distribution, which measure of central tendency is more resistant to outliers?",
    answer("Mean"),
    answer("Median", correct = TRUE),
    answer("Standard deviation"),
    answer("Variance"),
    incorrect = "The median is more resistant to outliers in a skewed distribution.",
    allow_retry = TRUE
  )
)
```




## Submit Report

```{r context="setup"}
submission_ui
```

```{r context="server"}
submission_server()
```
