---
title: "SSPS [4102|6006] Week 12: Network and Spatial data"
tutorial:
  id: "week-12"
output:
  learnr::tutorial:
    progressive: true
    ace_theme: github
    theme: united
runtime: shiny_prerendered
description: "This tutorial will cover machine learning and text analysis"
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(sspslearnr)
library(gradethis)
library(class)
library(tidytext)

tutorial_options(exercise.checker = gradethis::grade_learnr, 
                 exercise.reveal_solution="TRUE")
knitr::opts_chunk$set(echo = FALSE)
num_tutorial <- 0
num_excercise <- 0
tut_reptitle <- "SSPS [4102|6006] Week 12: In-class Individual Tutorial"
```

## 1. Import the data

```{r import-quiz}
quiz(
  question("Which function is most commonly used to load data from a CSV file into R?",
           answer("read.csv()", correct = TRUE),
           answer("load.file()"),
           answer("import.data()"),
           allow_retry = TRUE,
           correct = "Correct! The `read.csv()` function is typically used to load data from a CSV file into R.",
           incorrect = "Not quite. The correct function is `read.csv()`."
  ),
  
  question("What is the easiest way to import a dataset in RStudio without writing code?",
           answer("Using the Import Dataset button in the Environment pane", correct = TRUE),
           answer("Using the `read.import()` function"),
           answer("Using the 'Load Data' command"),
           allow_retry = TRUE,
           correct = "Correct! RStudio provides a graphical interface to import datasets via the Import Dataset button in the Environment pane.",
           incorrect = "Not quite. The Import Dataset button in RStudio’s Environment pane is the easiest way to load data without code."
  )
)
```

- Now complete the code so to import a CSV file "beauty.csv" from the folder "data" into an object R data.frame called `beauty` using the `read.csv()` function.

```{r importcode, exercise = TRUE, exercise.blanks = "___+"}

```

```{r importcode-solution}
beauty <- read.csv("data/beauty.csv")
```

```{r importcode-check}
grade_code()
```

## 2. Data exploration

- Now that we have loaded the data frame `beauty`, do a summary of all its columns with the function `summary()`. 

```{r load-beauty}
beauty <- 
  read.csv("data/beauty.csv")
```

```{r summary, exercise = TRUE, exercise.blanks = "___+", exercise.setup = 'load-beauty'}

```

```{r summary-solution}
summary(beauty)
```

```{r summary-check}
grade_code()
```

It is always important to know your data! The table contains data from anonymous teaching evaluations in economics classes at the University of Texas. 

This is real data collected for this study if you want to read it: Hamermesh, D. S., & Parker, A. (2005). Beauty in the classroom: Instructors’ pulchritude and putative pedagogical productivity. *Economics of Education Review*, 24(4), 369–376. https://doi.org/10.1016/j.econedurev.2004.07.013

The exercise was instead published in Gelman, A., Hill, J., & Vehtari, A. (2021). *Regression and Other Stories*. Cambridge University Press.

These are the variable:

- `eval`: The student evaluation on a 1-5 Likert scale.
- `beauty`: The instructor beauty according to ratings from six undergraduate students (three women and three men) based on the instructor's picture. This variable was standardised. 
- `female`: Whether the instructor is a female.
- `age`: Age of instructor.
- `minority`: Whether the instructor is part of a minority.
- `nonenglish`: Whether the instructor is not a native English speaker.
- `lower`: Whether the course was offered in the first years of the program.
- `course_id` The ID identifying the course.

You can come back to this tab to check what each variable means.

## 3. Data description

Let's now describe the key variable from the data: `eval`.

```{r description1-quiz}
quiz(
  question("Using ggplot, which graph can you use to show the distribution of a single continuous variable?",
           answer("geom_line()"),
           answer("geom_density()", correct = TRUE),
           answer("geom_bar()"),
           answer("geom_histogram()", correct = TRUE),
           answer("geom_col()"),
           answer("geom_boxplot()"),
           allow_retry = TRUE,
           correct = "Correct! With a single continous variable you can use both `geom_density()` or geom_histogram(). If you consider the Likert scale as a discrete variable you could use `geom_col()` or `geom_bar()`.",
           incorrect = "Not quite."
  )
)
```

### Now let's plot the variable `eval` with ggplot.

```{r plot1, exercise = TRUE, exercise.blanks = "___+", exercise.setup = 'load-beauty'}

```

<div id="plot1-hint">
**Hint:** You might want to review this slide from Week 03: https://fraba.github.io/SSPS4102-SSPS6006/slide/week-03.html#55.
</div>

```{r plot1-solution}
ggplot(data = beauty, aes(x = eval)) + 
  geom_histogram()
```

```{r plot1-check}
grade_code()
```

### It might also be useful to provide some statistics about your key variables.

What is the mean of the variable `eval`. 

```{r mean, exercise = TRUE, exercise.blanks = "___+", exercise.setup = 'load-beauty'}

```

<div id="mean-hint">
**Hint:** Remember that you can access a variable in a data frame using the `$` after the data frame name.
</div>

```{r mean-solution}
ggplot(data = beauty, aes(x = eval)) + 
  geom_histogram()
```

```{r mean-check}
mean(beauty$eval)
```

### Finally, it could be useful to visualise the associations among key variables.

- Can you plot with `geom_point()` the association between `eval` and `beauty`?

```{r plot2, exercise = TRUE, exercise.blanks = "___+", exercise.setup = 'load-beauty'}

```

<div id="plot2-hint">
**Hint:** You might want to review this slide from Week 03: https://fraba.github.io/SSPS4102-SSPS6006/slide/week-03.html#64.
</div>

```{r plot2-solution}
ggplot(data = beauty, aes(x = eval, beauty)) + 
  geom_point()
```

```{r plot2-check}
grade_code()
```

## 4. Data Manipulation

Likely, you are not going to use all the variables you find in a data set. Also, you might need to **filter** your data set (excluding observations you don't need) and **recode** some of the variables (maybe creating a new variable). 

Let's review how to do this with dplyr.

### Create a new data frame called `beauty_nonenglish` where you have only non-English instructors

```{r filter, exercise = TRUE, exercise.blanks = "___+", exercise.setup = 'load-beauty'}
beauty_nonenglish <- 
  beauty |>
  ___(nonenglish ___ ___)
```

<div id="filter-hint">
**Hint:** To affect rows, you might want to review the slides from Week 05: https://fraba.github.io/SSPS4102-SSPS6006/slide/week-05.html#31.
</div>

```{r filter-solution}
beauty_nonenglish <- 
  beauty |>
  filter(nonenglish == 1)
```

```{r filter-check}
grade_code()
```


### Use mutate(), case_when() and pipes to create a variable called `age_group` that groups age into thre categories: "21-30", "31-40", "41+" (use this also as labels of your categories).

```{r casewhen, exercise = TRUE, exercise.blanks = "___+", exercise.setup = 'load-beauty'}
beauty <- 
  beauty |>
  ___(___ = case_when(age >= ___ & age <=___ ~ "21-30",
                      age >= ___ & age <= ___ ~ "31-40",
                      age >= ___ ~ "41+"))
```

<div id="casewhen-hint">
**Hint:** To transform columns, you might want to review the slides from Week 05: https://fraba.github.io/SSPS4102-SSPS6006/slide/week-05.html#42.
</div>

```{r casewhen-solution}
beauty <- 
  beauty |>
  mutate(age_group = case_when(age >= 21 & age <= 30 ~ "21-30",
                               age >= 31 & age <= 40 ~ "31-40",
                               age >= 41 ~ "41+"))
```

```{r casewhen-check}
grade_code()
```

## 5. Data analysis

Once you have read in and prepared your data you can run your analysis. 

### Let's start simple and only analysie the relationship between two variable.

- Use the function `cor()` to measure the *correlation* between `eval` and `beauty`. 

```{r cor, exercise = TRUE, exercise.blanks = "___+", exercise.setup = 'load-beauty'}
cor(___)
```

<div id="cor-hint">
**Hint:** On this you can check the stat slides from Week 06: https://canvas.sydney.edu.au/courses/59609/files/39052980?module_item_id=2474521.
</div>

```{r cor-solution}
cor(beauty$eval, beauty$beauty)
```

```{r cor-check}
grade_code()
```

```{r cor-quiz}
quiz(
  question("What does the correlation coefficient (cor(X,Y)) represent in data analysis?",
           answer("The exact cause of variation between two variables"),
           answer("The direction and strength of the linear relationship between two variables", correct = TRUE),
           answer("The difference between two variables"),
           allow_retry = TRUE,
           correct = "Correct! The correlation coefficient measures the direction and strength of the linear relationship between two variables.",
           incorrect = "Not quite. The correlation coefficient represents the direction and strength of the linear relationship between two variables."
  ),
  
  question("What is the possible range of values for the correlation coefficient?",
           answer("0 to 1"),
           answer("0 to 100"),
           answer("-1 to 1", correct = TRUE),
           allow_retry = TRUE,
           correct = "Correct! The correlation coefficient ranges from -1 (perfect negative correlation) to 1 (perfect positive correlation).",
           incorrect = "Not quite. The correct range is -1 to 1."
  ),
  
  question("If the correlation coefficient between two variables is close to 0, what does that imply?",
           answer("There is no strong linear relationship between the two variables", correct = TRUE),
           answer("The variables are not related in any way"),
           answer("One variable directly causes the other"),
           allow_retry = TRUE,
           correct = "Correct! A correlation close to 0 suggests there is no strong linear relationship between the variables.",
           incorrect = "Not quite. A correlation close to 0 means there is no strong *linear* relationship, but they might still be related in other ways."
  )
)
```

### Let's use the linear regression to predict `eval`.

```{r predict-quiz}
quiz(
  question("When we say that we want to predict a variable (e.g., `eval`)...",
           answer("We consider that variable the outcome variable", correct = TRUE),
           answer("We consider that variable a control variable"),
           answer("We consider that variable the dependent variable", correct = TRUE),
           answer("We consider that variable the independent variable"),
           allow_retry = TRUE,
           correct = "Correct! Outcomes and dependent variable is what we want to predict with a regression.",
           incorrect = "Not quite."
  )
)
```

- Write your code to predict `eval` with a linear regression `lm()` using `beauty` as the only predictor.

```{r lm, exercise = TRUE, exercise.blanks = "___+", exercise.setup = 'load-beauty'}
lm(formula = ___, data = beauty)
```

<div id="lm-hint">
**Hint:** Remember that the formula for the regression is `Y ~ X` where `Y` is the dependent variable and `X` are all the independent variables.  
</div>

```{r lm-solution}
lm(formula = eval ~ beauty, data = beauty)
```

```{r lm-check}
grade_code()
```


```{r coeff-1}
question_numeric(
  "What is the coefficient of `beauty`?", 
  answer(0.133, correct = TRUE), 
  allow_retry = TRUE, min = 0.132, max = 0.134, step = 0.01 )
```

### Let's complicate things a bit. 

```{r control-quiz}
quiz(
  question("What is a confounding variable in the context of observational data analysis?",
           answer("A variable that only affects the outcome"),
           answer("A variable that affects both the treatment and the outcome", correct = TRUE),
           answer("A variable that only affects the treatment"),
           allow_retry = TRUE,
           correct = "Correct! A confounding variable influences both the likelihood of receiving the treatment and the outcome.",
           incorrect = "Not quite. A confounding variable affects both the treatment and the outcome."
  ),
  
  question("Why do we include control variables in a multiple regression model?",
           answer("To increase the sample size"),
           answer("To account for potential confounders and isolate the effect of the treatment on the outcome", correct = TRUE),
           answer("To change the relationship between the treatment and the outcome"),
           allow_retry = TRUE,
           correct = "Correct! Control variables help to account for potential confounders, allowing us to estimate the treatment effect more accurately.",
           incorrect = "Not quite. Control variables are used to account for confounders, not to change the relationship."
  ),
  
  question("What is the main assumption we make when controlling for confounders using multiple regression?",
           answer("We assume that confounding variables do not affect the outcome"),
           answer("We assume that we have included all relevant confounding variables in the model", correct = TRUE),
           answer("We assume that confounders are independent of the treatment"),
           allow_retry = TRUE,
           correct = "Correct! When using multiple regression, we assume that all relevant confounding variables are included in the model to estimate the treatment effect accurately.",
           incorrect = "Not quite. We need to assume that all confounders are accounted for in the model."
  )
)
```


- Let's add a control variable to your linear model. How would you control for `age` while predicting `eval` with `beauty` (this time we store the result in `fit`)?


```{r lm2, exercise = TRUE, exercise.blanks = "___+", exercise.setup = 'load-beauty'}
fit <- lm(formula = ___, data = beauty)
```

<div id="lm2-hint">
**Hint:** Remember that you can add control variables to your formula with a `+` after your independent variable.  
</div>

```{r lm2-solution}
fit <- lm(formula = eval ~ beauty + age, data = beauty)
```

```{r lm2-check}
grade_code()
```

- Let's check the results from the regression with `summary()` (note that they are stored in `fit`).

```{r beautylm}
beauty <- read.csv("data/beauty.csv")
fit <- lm(formula = eval ~ beauty + age, data = beauty)
```

```{r lm3, exercise = TRUE, exercise.blanks = "___+", exercise.setup = 'beautylm'}
___
```


```{r coeff-2}
question_numeric(
  "What is the coefficient of `beauty` now?", 
  answer(0.1340634, correct = TRUE), 
  allow_retry = TRUE, min = 0.134, max = 0.135, step = 0.01 )
```

```{r coeff-quiz}
quiz(
  question("Which of the following is the correct interpretation of a coefficient in a multiple linear regression model?",
           answer("It represents the change in the outcome variable for a one-unit increase in the predictor variable, without considering other variables"),
           answer("It represents the change in the predicted value of the outcome variable for a one-unit increase in the predictor variable, holding all other variables constant", correct = TRUE),
           answer("It represents the percentage change in the outcome variable for a one-unit increase in the predictor variable"),
           allow_retry = TRUE,
           correct = "Correct! The coefficient reflects the change in the predicted outcome for a one-unit increase in the predictor, holding other variables constant.",
           incorrect = "Not quite. Remember that the coefficient shows the effect of a predictor variable while controlling for other variables in the model."
  )
)
```

```{r pvalue-quiz}
quiz(
  question("Is the coefficient of `beauty` statistically significant at the .05 level?",
           answer("Yes", correct = TRUE),
           answer("No")
  ), 
  question("Which of the following is the correct interpretation of a p-value in hypothesis testing?",
           answer("It is the probability of observing data as extreme as the observed data, assuming the null hypothesis is true", correct = TRUE),
           answer("It is the probability that the null hypothesis is true"),
           answer("It is the probability that the alternative hypothesis is true"),
           allow_retry = TRUE,
           correct = "Correct! The p-value represents the likelihood of observing data as extreme as the observed data under the null hypothesis.",
           incorrect = "Not quite. The p-value reflects the probability of the observed data under the assumption that the null hypothesis is true."
  )
)
```

- What other control variables would you add to predict `eval` from `beauty`? Why?

```{r lm4, exercise = TRUE, exercise.blanks = "___+", exercise.setup = 'load-beauty'}
fit <- lm(formula = ___, data = beauty)
summary(fit)
```


## Submit Report

```{r context="setup"}
submission_ui
```

```{r context="server"}
submission_server()
```


