---
title: "SSPS [4102|6006] Week 07: Probability"
tutorial:
  id: "week-07"
output:
  learnr::tutorial:
    progressive: true
    ace_theme: github
    theme: united
runtime: shiny_prerendered
description: "This tutorial will cover `lm()`, `predict()` and basic probability concepts. (Questions from tutorial designed by Matt Blackwell)"
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(sspslearnr)
library(nycflights13)
library(gradethis)

tutorial_options(exercise.checker = gradethis::grade_learnr, 
                 exercise.reveal_solution="TRUE")
knitr::opts_chunk$set(echo = TRUE)
num_tutorial <- 0
num_excercise <- 0
tut_reptitle <- "SSPS [4102|6006] Week 07: In-class Individual Tutorial"
```


## `lm()`

Using the data set `mtcars`, run a regression predicting `mpg` using as predictors `wt` and `hp`. 

```{r lm-1, exercise = TRUE}

```

```{r lm-1-solution}
fit <- lm(mpg ~ wt + hp,  data = mtcars)
summary(fit)
```

Then report below $\hat{\beta}$ for the variable `wt` (round it to two digits).

```{r lm-2, echo = FALSE}
question_text(
  "Answer",
  answer("-3.88", correct = TRUE),
  answer("-3.87", correct = TRUE),
  allow_retry = TRUE
)
```

### Conceptual questions

```{r lm-3, echo = FALSE}
quiz(caption = "",
     question("What does the coefficient of a variable in a multiple linear regression model represent?",
              answer("The average effect of a one-unit change in that variable, holding other variables constant", correct = TRUE),
              answer("The total change in the dependent variable"),
              answer("The predicted value of the dependent variable")),
              
     question("If a coefficient is very small, what does this imply?",
              answer("The variable has a measurable but weak association with the dependent variable", correct = TRUE),
              answer("The variable has no association with the dependent variable"),
              answer("The model is invalid")),
              
     question("What is the difference between multiple R-squared and adjusted R-squared?",
              answer("Adjusted R-squared accounts for the number of predictors in the model", correct = TRUE),
              answer("Multiple R-squared always gives a higher value than adjusted R-squared"),
              answer("Multiple R-squared penalises for adding more predictors"))
     )
```


## `lm() with categorical variables as predictors`

The `iris` data frame contains observations of 50 flowers from 3 species of iris.

```{r}
head(iris)
```

Fit the `iris` data into a model that predicts `Sepal.Length` using the variables `Petal.Length`, `Petal.Width` and `Species`, then print key model summary statistics with `summary()`.

```{r predict-1, exercise = TRUE}

```

```{r predict-1-solution}
fit <- lm(Sepal.Length ~ Petal.Length + Petal.Width + Species, data = iris)
summary(fit)
```

### Wait a minute!

The variable `Species` is a categorical variable with three levels (`r paste0(levels(iris$Species), collapse = ", ")`)! Yes, but the interpretation of the coefficient is only slightly different.

In this case, the coefficient `Speciesversicolor` (i.e. the versicolor species) represents how much the predicted outcome (`Sepal.Length`) differs for the versicolor species compared to the reference species (in this case the setosa species).

## predict()

Now use this model to predict the sepal length for a flower of the species virginica, with petal length 4 and petal width 2.

```{r predict-2, exercise = TRUE}
fit <- lm(Sepal.Length ~ Petal.Length + Petal.Width + Species, data = iris)
# Add your predict(fit, newdata = ...) code here
```

```{r predict-2-solution}
fit <- lm(Sepal.Length ~ Petal.Length + Petal.Width + Species, data = iris)
predict(fit, newdata = data.frame(Petal.Length = 4, Petal.Width = 2, Species = "virginica"))
```

Enter your prediction for the sepal length below (round to two digits)

```{r predict-3, echo = FALSE}
question_text(
  "Answer",
  answer("5.18", correct = TRUE),
  allow_retry = TRUE
)
```

## Probability Basics

### Conceptual questions

```{r concept, echo=FALSE}
quiz(
  caption = "",
  question(
    "I roll a fair 6-sided die 3,000,000 times. The number 1 turns up approximately 500,000 times, so I conclude that the probability of a 1 is 1/6. What approach to probability does this ",
    answer("Frequentist", correct = TRUE),
    answer("Bayesian"))
)
```



### Bounding probabilities

Suppose that P(A)=0.4 and P(B)=0.3. What is the minimum value of P(A and not B)?


```{r ltp, echo = FALSE}
question_text(
  "Answer:",
  answer("0.1", correct = TRUE),
  answer(".1", correct = TRUE),
  allow_retry = TRUE
)
```


<div id="filter-hint">
**Hint:** See the Law of Total Probability. What happens if B is entirely contained in A?
</div>


### Combinations/Permutations

```{r combn, echo = FALSE}
question("Out of all possible anagrams for the word CHAIR (including the original word), what is the probability that the letter I comes first?",
         answer("0.1"),
         answer("0.2", correct = TRUE),
         answer("0.3"),
         answer("0.4"))
```


### Independence

Let A and B be two independent variables that take on only the values 0 and 1. If P(A=1, B=1)=0.12, P(A=0, B=1)=0.28, and P(A=1)=0.3, what is P(A=0, B=0)? Use this code block to perform calculations and enter the probability below.


```{r indep_chunk, exercise = TRUE}

```


```{r indep, echo = FALSE}
question_text(
  "Answer:",
  answer("0.42", correct = TRUE),
  answer(".42", correct = TRUE),
  allow_retry = TRUE
)
```


## Submit Report

```{r context="setup"}
submission_ui
```

```{r context="server"}
submission_server()
```
