---
params:
  student_name: "Student T."
  student_unikey: "abcd1234"
  reportdate: !r Sys.Date()
  reporttitle: "Week 0 Tutorial Report"
  skipped: 0
  checked: 0
  output: !r tibble::tibble(id = "test")
title: "`r params$reporttitle`"
date: "`r params$reportdate`"
---

* STUDENT Name: `r params$student_name`

* STUDENT UniKey: `r params$student_unikey`

```{r, echo = TRUE}
library("sspslearnr")

print(params)

checked_prop <- sum(params$output$checked, na.rm = T) / length(params$output$checked)
```

Congrats on getting through this tutorial! Please upload this PDF to Canvas/Gradescope. 

## Tutorial progress:

- Exercises attempted: `r length(params$output$attempted)`
- Exercises attempted (%): `r round(checked_prop*100,2)`
- Correct exercises: `r sum(params$output$checked, na.rm = T) `

\newpage

## Progress detail

Here is a detailed view of your progress

```{r, echo = FALSE, center = TRUE}
knitr::kable(params$output[, c("id", "attempted")], "latex")
```

\newpage

## Environment

```{r, echo = TRUE}
R.version
Sys.time()
Sys.info()[c("user", "effective_user")]
```
