---
title: "SSPS [4102|6006] Week 08: Uncertainty"
tutorial:
  id: "week-08"
output:
  learnr::tutorial:
    progressive: true
    ace_theme: github
    theme: united
runtime: shiny_prerendered
description: "This tutorial will cover `lm()` and hypothesis testing."
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(sspslearnr)
library(gradethis)

tutorial_options(exercise.checker = gradethis::grade_learnr, 
                 exercise.reveal_solution="TRUE")
knitr::opts_chunk$set(echo = FALSE)
num_tutorial <- 0
num_excercise <- 0
tut_reptitle <- "SSPS [4102|6006] Week 08: In-class Individual Tutorial"
```


## Quiz: Understanding the Null Hypothesis

```{r null_hypothesis_quiz}
question("What is the null hypothesis in hypothesis testing?",
         answer("The hypothesis that the true effect is non-zero."),
         answer("The hypothesis that the sample mean equals the population mean."),
         answer("The hypothesis that the true effect is zero.", correct = TRUE),
         allow_retry = TRUE,
         correct = "Correct! The null hypothesis usually states that there is no effect or relationship.",
         incorrect = "Not quite. The null hypothesis is typically about no effect or no relationship."
)

```

## Calculate the average causal effect of attending a small class on math test scores using the provided data

```{r estimate_ate, exercise = TRUE, exercise.blanks = "___+"}
# Load the data (don't change this)
star <- read.csv("data/STAR.csv")

# Create the treatment variable
star$small <- ifelse(star$classtype == "small", 1, 0)

# Now on the line below fit a linear model with lm() predicting
# the `math` test result using the treatment variable `small`
fit <- lm(____)

# Finally, display the summary of the model
summary(fit)
```

```{r estimate_ate-solution}
star <- read.csv("STAR.csv")
star$small <- ifelse(star$classtype == "small", 1, 0)
fit <- lm(star$math ~ star$small)
summary(fit)
```

### What is the average causal effect of attending a small class on math test scores? Round all values to two decimal places.

```{r estimate_ate-1}
question_text(
  "Answer",
  answer("5.99", correct = TRUE),
  answer("5.98", correct = TRUE),
  allow_retry = TRUE
)
```

###  What is the test-statistic for $\hat{\beta}$ Round all values to two decimal places.

```{r estimate_ate-2}
question_text(
  "Answer",
  answer("2.75", correct = TRUE),
  allow_retry = TRUE
)
```

###  Interpreting Regression Coefficients

```{r interpret_coefficients}
question("What does the coefficient for 'star$small' represent in the regression output?",
  answer("The estimated average difference in math scores between students in small and regular classes.", correct = TRUE),
  answer("The estimated average math score for all students."),
  answer("The estimated average difference in reading scores."),
  allow_retry = TRUE,
  correct = "Correct! The coefficient represents the average difference in scores between the two groups.",
  incorrect = "Incorrect. Remember, the coefficient directly measures the effect of the treatment variable."
)
```

###  Conclusion on Statistical Significance

```{r conclusion_significance}
question("Based on the calculated p-value, is the effect of attending a small class statistically significant at the 5% level?",
  answer("Yes, because the p-value is less than 0.05.", correct = TRUE),
  answer("No, because the p-value is greater than 0.05."),
  allow_retry = TRUE,
  correct = "Correct! A p-value less than 0.05 indicates that the effect is statistically significant.",
  incorrect = "Incorrect. Check the calculated p-value and compare it to the 5% threshold."
)
```

## Calculating the test-statistic of a regression coefficient

OK, this is a bit hard as it requires remembering the formula for the test-statistic...

### First, estimate the average causal effect of attending a small class on the reading test scores?

```{r estimate_ate-3, exercise = TRUE, exercise.blanks = "___+"}
# Load the data (don't change this)
star <- read.csv("data/STAR.csv")

# Create the treatment variable
star$small <- ifelse(star$classtype == "small", 1, 0)

# Now on the line below fit a linear model with lm() predicting
# the `reading` test result using the treatment variable `small`
fit <- lm(____)

# Finally, display the summary of the model
____
```

```{r estimate_ate-3-solution}
star <- read.csv("data/STAR.csv")
star$small <- ifelse(star$classtype == "small", 1, 0)
fit <- lm(star$reading ~ star$small)
summary(fit)
```

### What is the p-value of the `reading` coefficient? Round the p-value to four decimal places.

```{r estimate_ate-4}
question_text(
  "Answer",
  answer("0.0004", correct = TRUE),
  answer("0.0005", correct = TRUE),
  allow_retry = TRUE
)
```

```{r conclusion_significance-2}
question("Based on the calculated p-value, is the effect of attending a small class on the reading test statistically significant at the 5% level?",
  answer("Yes", correct = TRUE),
  answer("No"),
  allow_retry = TRUE,
  correct = "Correct! A p-value less than 0.05 indicates that the effect is statistically significant.",
  incorrect = "Incorrect. Check the calculated p-value and compare it to the 5% threshold."
)
```

### Now, using the values you can find in the summary of the model, calculate the test-statistic for the coefficient of `small` and store it in the variable `z_obs`. Round all values to two decimal places. 

```{r z-obs, exercise = TRUE, exercise.blanks = "___+"}
# Test statistic
z_obs <- ___ / ___
```

```{r z-obs-solution}
# Test statistic
z_obs <- 7.21 / 2.06
```

```{r z-obs-check}
grade_code()
```

## Submit Report

```{r context="setup"}
submission_ui
```

```{r context="server"}
submission_server()
```
