---
title: "SSPS [4102|6006] Week 06: Prediction"
tutorial:
  id: "week-06"
output:
  learnr::tutorial:
    progressive: true
    ace_theme: github
    theme: united
runtime: shiny_prerendered
description: "This tutorial will cover basics of regression analysis with two variables."
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(sspslearnr)
library(nycflights13)
library(gradethis)

tutorial_options(exercise.checker = gradethis::grade_learnr, 
                 exercise.reveal_solution="TRUE")
knitr::opts_chunk$set(echo = TRUE)
num_tutorial <- 0
num_excercise <- 0
tut_reptitle <- "SSPS [4102|6006] Week 06: In-class Individual Tutorial"

# Data
set.seed(12345)
n <- 10
x <- sample(100)
y <- x + rnorm(100, mean = 0, sd = 0.5)
```


## Conceptual questions

```{r lm-multi, echo = FALSE}
quiz(
  caption = "",
  question(
    "We use the regression line to predict the value of the outcome variable y hat, otherwise known as:",
    answer("predicted value"),
    answer("fitted value"),
    answer("both of these", correct = TRUE)),
  question(
    "The differences between the observed outcome and its predicted value is called a:",
    answer("residual"),
    answer("prediction error"),
    answer("either of these", correct = TRUE)),
  question(
    "An R^2 close to 1 suggests the model fits:",
    answer("well", correct = TRUE),
    answer("not well")),
  question(
    "With a single binary independent variable, what is the interpretation of the slope coefficient in a linear regression model?",
    answer("Average outcome in the `X = 1` group"),
    answer("Average outcome in the `X = 0` group"),
    answer("Differences in means between the `X=1` and `X=0` groups", correct = TRUE))
)
```

## Interpreting intercepts

In the following box, type answer to the following question: The intercept in the model $Y_i = \alpha + \beta X_i + \epsilon_i$ represents the average value of $Y_i$ when $X_i$ is equal to what?

```{r intercept, echo = FALSE}
question_text(
  "The intercept is the average of Y when X is equal to:",
  answer("0", correct = TRUE)
)
```




## Running regressions in R

We have included two vectors `x` and `y`. Regress `y` on `x` and assign it to an object called `fit` and print that object to see what the output looks like.

```{r lm, exercise = TRUE}

```

```{r lm-solution}
fit <- lm(y ~ x)
fit
```

```{r lm-check}
grade_result(
  pass_if(~ identical(coef(.result), coef(lm(y~x))))
)
```

<div id="lm-hint">
**Hint:** To print an object just add its name to your code.
</div>

## Accessing output from lm

Now, use the `coef()` function to access the two coefficients in the `fit` object.

```{r coef, exercise = TRUE}
fit <- lm(y ~ x)

```

```{r coef-check}
grade_result(
  pass_if(~ identical(.result, coef(fit)), "Good job!")
)
```


Now you can also access the fitted values using the `fitted()` function. Print the fitted values here:

```{r fitted, exercise = TRUE}
fit <- lm(y ~ x)

```

```{r fitted-check}
grade_result(
  pass_if(~ identical(.result, fitted(fit)) | identical(.result, predict(fit)),
          "Good job!")
)
```

Do you understand what the values from `predict()` represent? 

```{r lm-quiz, echo = FALSE}
quiz(
  question(
    "What does values does `predict(fit)` return?",
    answer("The residuals from the linear model, which represent the difference between the observed and predicted y values."),
    answer("The estimated coefficients (slope and intercept) from the linear model."),
    answer("The predicted `y` values for the `x` values used in the model.", correct = TRUE),
    answer("A list of statistical metrics such as R-squared, p-values, and standard errors.")
  )
)
```


## Model fit

Use the `summary()` function to access the value of $R^2$, which is named `Multiple R-squared` in summary output, and replace with its values the `#` in the code.

```{r rsq, exercise = TRUE}
fit <- lm(y ~ x)
summary(fit)
#
```


```{r rsq-check}
grade_result(
  pass_if(~ identical(.result, 0.9997), "Good job!")
)
```

###

Yes, if you have two variables you can also calculate your $R^2$ with `cor(y, x)^2` and you will get the exact same result.

```{r cor-1, exercise = TRUE}
cor(y, x)^2
```

## Using `lm()`

Now, consider the `mtcars` dataset.

```{r mtcars-1, exercise = TRUE}
head(mtcars)
```

Using the `mtcars` data and replacing `...`, fit a linear model with `lm()` where where `mpg` is the dependent variable, and `hp` is the independent variable.

```{r mtcars-2, exercise = TRUE}
fit <- ...
fit
```

```{r mtcars-2-solution}
fit <- lm(mpg~hp, data = mtcars)
fit
```

```{r mtcars-2-check}
grade_result(
  pass_if(~ identical(coef(.result), coef(lm(mpg~hp, data = mtcars))), "Good job!")
)
```

Using `mtcars`, model the relation between `wt` (the weight of the car in 1000 lbs) as your independent variable and `mpg` (miles per gallon) as dependent variable.

```{r mtcars-3, exercise = TRUE}
fit <- ...
fit
```

```{r mtcars-3-solution}
fit <- lm(mpg~wt, data = mtcars)
fit
```

```{r mtcars-3-check}
grade_result(
  pass_if(~ identical(coef(.result), coef(lm(mpg~wt, data = mtcars))), "Great")
)
```

Based on linear model you just estimated with `lm()`, use `predict()` to find the predicted value or outcome ($\hat{Y}$) for the "Ferrari Dino". Enter the value below replacing `#____`). ("Run" `lm()` and `predict()` in the first block to find the value.)

```{r mtcars-4-0, exercise = TRUE}
fit <- lm(mpg~wt, data = mtcars)
predict(fit)
```

```{r mtcars-4, exercise = TRUE}
pred_ferrari_dino <- #____
pred_ferrari_dino
```


```{r mtcars-4-solution}
predict(fit)
pred_ferrari_dino <- 22.480940
pred_ferrari_dino
```

```{r mtcars-4-check}
grade_result(
  pass_if(~ identical(.result, 22.480940), "Great")
)
```

Now, how do you calculate the prediction error $\hat{\epsilon}$ (as in $\hat{\epsilon} = Y - \hat{Y}$) for the "Ferrari Dino"? 

First, you will need to find the actual observed value for your dependent variable $Y$ for that car. You can do this (since the car name is also the row name ) ...

```{r mtcars-5, exercise = TRUE }
mtcars["Ferrari Dino",]
```

Then you can compute the error  $\hat{\epsilon}$ (as in $\hat{\epsilon} = Y - \hat{Y}$) (use the predicted value $\hat{Y}$ you estimated and visualised above).

```{r mtcars-6, exercise = TRUE, exercise.blanks = "___+"}
error <- ____ - ____
error
```

```{r mtcars-6-solution}
error <- 19.7 - 22.480940
error
```

```{r mtcars-6-check}
grade_result(
  pass_if(~ identical(.result, 19.7 - 22.480940), "Yes!")
)
```

## Submit Report

```{r context="setup"}
submission_ui
```

```{r context="server"}
submission_server()
```
