---
title: "SSPS [4102|6006] Week 03: Data visualisation"
tutorial:
  id: "week-03"
output:
  learnr::tutorial:
    progressive: true
    ace_theme: github
    theme: united
runtime: shiny_prerendered
description: "This tutorial will cover data visulisation with ggplot2."
---

```{r setup, include=FALSE}
library(gradethis)
library(learnr)
library(sspslearnr)
library(ggplot2)

tutorial_options(exercise.checker = gradethis::grade_learnr, 
                 exercise.reveal_solution="TRUE")
knitr::opts_chunk$set(echo = FALSE)
num_tutorial <- 0
num_excercise <- 0
tut_reptitle <- "SSPS [4102|6006] Week 03: In-class Individual Tutorial"

## Data
gapminder <-
  readRDS("data/gapminder.rds")

```

**NOTE: Make sure to click on each "Run code" sequentially (from top to bottom) even if there's no question or you are not asked to modify the R code.**

## Setup

Make sure you have installed ggplot2 on your computer. Try this,

```{r load-ggplot, exercise = TRUE}
library(ggplot2)
```

* If you get no error, you are good to go! The package ggplot2 was successfully loaded.

* If you get "Error in library(ggplot2) : there is no package called ‘ggplot2’" then make sure you install the package with

```{r eval = FALSE, echo = TRUE}
install.packages("ggplot2")
# Then load it with
library(ggplot2)
```


## About `ggplot2`

A plot in `ggplot2` consists of several layers:

1. data
    - strict format: tidy data!
    - one row per datapoint
    - all grouping must be included in the data

2. aesthetics (`aes`) ([original documentation](https://ggplot2.tidyverse.org/reference/index.html#section-aesthetics))
    - how to map the data to the graph?
    - which column is *x*, *y*, ...
    - which column provides grouping (based on the grouping, one can either connect points into lines, split a graph into several facets, or color differently each group)

3. `geom`s ([original documentation](https://ggplot2.tidyverse.org/reference/index.html#section-geoms))
    - how to visualize the data?
    - points (`geom_point`), lines (`geom_line`), bars (`geom_bar`), etc.
    - `geom`s are connected to `stat`s that conduct any necessary pre-processing of data (e.g., `geom_histogram` would first calculate the number of observations in each bin through `stat_bin`)

4. `scale`s ([original documentation](https://ggplot2.tidyverse.org/reference/index.html#section-scales))
    - any type of data representation on the plot
    - coordinates: `scale_x_continuous`, `scale_x_discrete`, `scale_x_date`, etc.
    - colors: `scale_color_manual`, `scale_colour_brewer`, `scale_fill_continuous`, etc.
    
Each layer is added to the previous one with a `+` sign. The result can be saved to an object and added upon later. 

(Credit for this explanation: Julia Romanowska)


## Gapminder data

Let's consider the [Gapminder](https://cran.r-project.org/web/packages/gapminder/readme/README.html) data set.

```{r}
gapminder
```


```{r gapminder-quiz-1, echo=FALSE}
quiz(
  question("What does each row in the dataset represent?",
           answer("A specific country's statistics for a given year", correct = TRUE),
           answer("A continent's yearly summary", correct = FALSE),
           answer("A worldwide summary for a specific year", correct = FALSE),
           answer("Statistics of multiple countries combined", correct = FALSE),
           allow_retry = TRUE))
```

---

## First visualisation

Points are probably the simplest way to visualise two continuous variables. Run the following code

```{r plot-1-1, exercise = TRUE}
ggplot(data = gapminder) + 
  geom_point(mapping = aes(x = lifeExp, y = gdpPercap))
```

```{r gapminder-quiz-2, echo=FALSE}
quiz(
  question("What does the `aes()` function do in the code?\n\n`ggplot(data = gapminder) + geom_point(mapping = aes(x = lifeExp, y = gdpPercap))`",
           answer("Specifies the aesthetic mappings, linking the x and y variables to `lifeExp` and `gdpPercap`, respectively", correct = TRUE),
           answer("Adds a geometric shape to the plot", correct = FALSE),
           answer("Filters the data based on the given conditions", correct = FALSE),
           answer("Specifies the title of the plot", correct = FALSE),
           allow_retry = TRUE)
)
```

Now change the code so that changes in the `lifeExp` variables are mapped on the vertical axis while `gdpPercap` on the horizontal axis.

```{r plot-1-2, exercise = TRUE}
ggplot(data = gapminder) + 
  geom_point(mapping = aes(x = lifeExp, y = gdpPercap))
```

```{r plot-1-2-solution}
ggplot(data = gapminder) + 
  geom_point(mapping = aes(x = gdpPercap, y = lifeExp))
```

```{r plot-1-2-check}
grade_code()
```

## Add coloring 

Here is the magic of ggplot - we want to add coloring of the points *by one of
the variables in a dataset*. Therefore, we need to specify aesthetics
for the `geom_point`.

Let's consider the `mpg` data frame, where each row is a car model (`mpg` contains observations collected by the US Environmental Protection Agency on 234 models of car.)

```{r}
mpg
```

and let's consider this plot

```{r echo = TRUE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

> NOTE: we can save the plots as objects and then add on to those if we want to modify the plot!

```{r mpg-quiz-1, echo=FALSE}
quiz(
  question("In the following `ggplot2` code, which variable determines the colour of the points?\n\n`ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = class))`",
           answer("class", correct = TRUE),
           answer("displ", correct = FALSE),
           answer("hwy", correct = FALSE),
           answer("mpg", correct = FALSE),
           allow_retry = TRUE
  )
)
```

Now let's go back to our `gapminder` dataset. Edit the code so to use the variable `continent` for colouring your points.

```{r plot-1-3, exercise = TRUE}
ggplot(data = gapminder) + 
  geom_point(mapping = aes(x = lifeExp, y = gdpPercap))
```

```{r plot-1-3-solution}
ggplot(data = gapminder) + 
  geom_point(mapping = aes(x = lifeExp, y = gdpPercap, colour = continent))
```

```{r plot-1-3-check}
grade_this_code(correct = "Well done!",
                incorrect = "You should set the aesthetic `colour` or `color` within `aes()` to the corresponding variable.")
```

## Size and Colour

Edit this code so that the size of each point is proportional to the value of `gdpPercap` and the colour to `lifeExp` (life expectancy).

```{r plot-1-4, exercise = TRUE}
ggplot(data = gapminder) + 
  geom_point(mapping = aes(x = lifeExp, y = gdpPercap))
```

```{r plot-1-4-solution}
ggplot(data = gapminder) + 
  geom_point(mapping = aes(x = lifeExp, y = gdpPercap, size = gdpPercap, colour = lifeExp))
```

```{r plot-1-4-check}
grade_this_code()
```

```{r gapminder-quiz-3, echo=FALSE}
quiz(
  question("In the previous code consider the type of data `lifeExp` and `continent` represent, why is the legend for the `colour` aesthetic different when you map `colour` to `lifeExp` compared to mapping `colour` to `continent`?",
           answer("The legend is different because `lifeExp` is a categorical variable, so each level of life expectancy gets a different colour.", correct = FALSE),
           answer("The legend is different because `continent` is a continuous variable, so the legend shows a gradient of colours representing different continents.", correct = FALSE),
           answer("The legend is different because `lifeExp` is a continuous variable, so the legend shows a gradient of colours representing different ranges of life expectancy.", correct = TRUE),
           answer("The legend is different because `continent` is a discrete variable, so each continent gets a different distinct colour.", correct = TRUE),
           allow_retry = TRUE
  )
)
```

Finally, let's colour all points in red using `colour = "red"`. How would you do it?

```{r plot-1-5, exercise = TRUE}
ggplot(data = gapminder) + 
  geom_point(mapping = aes(x = lifeExp, y = gdpPercap))
```

```{r plot-1-5-solution}
ggplot(data = gapminder) + 
  geom_point(mapping = aes(x = lifeExp, y = gdpPercap), colour = 'red')
```

```{r plot-1-5-check}
grade_this_code(incorrect = 'Remember to set colour `"red"` with quotes outside of `aes()`')
```

## One variable categorical (a.k.a., discrete)

Now let's consider again the car models dataset ...

```{r}
mpg
```

Let's say I want to visualise with a bar how many models for each `manufacturer` are present in the data frame. Using `geom_bar()`, how would you do it?

```{r plot-1-6, exercise = TRUE}
ggplot(mpg) + ...
```

```{r plot-1-6-solution}
ggplot(mpg) + geom_bar(aes(x = manufacturer))
```

```{r plot-1-6-check}
grade_this_code(incorrect = 'You should specify the aesthetics of `geom_bar()` with `x = manufacturer`.')
```


```{r mpg-quiz-2, echo=FALSE}
quiz(
  question("Why does the y-axis represent a count?",
           answer("The y-axis represents a count because the dataset includes a `count` variable that is automatically used.", correct = FALSE),
           answer("The y-axis represents a count because the `aes()` function explicitly mapped the `y` aesthetic to a count variable.", correct = FALSE),
           answer("The y-axis represents a count because `geom_bar()` by default counts the number of occurrences of each category on the x-axis.", correct = TRUE),
           answer("The y-axis represents a count because `manufacturer` is a continuous variable and `geom_bar()` sums these values.", correct = FALSE),
           allow_retry = TRUE
  )
)
```

## One variable continuous

Now let's say that instead of a single categorical variable (`manufacturer`) you want to visualise a single continuous variable such as `lifeExp` in the gapminder data frame.

```{r plot-1-7, exercise = TRUE}
ggplot(gapminder) + geom_...(aes(x = lifeExp))
```



```{r gapminder-quiz-4, echo=FALSE}
quiz(
  question("Which one of these options would you use? You can test your code before answering",
    answer("geom_bar()", correct = FALSE),
    answer("geom_density()", correct = TRUE),
    answer("geom_col()", correct = FALSE),
    answer("geom_histogram()", correct = TRUE),
    allow_retry = TRUE
  )
)
```

## Subsetting data frames

Before moving on, let's learn something more about base R. Last week we learned how to subset vectors based on some conditions using  `[]`. 

Remember that if vectors are single-dimension data objects, data frames are instead two-dimension data objects. 

If I create this vector (please run the code)

```{r subset-1-1, exercise=TRUE}
my_vector <- c(1, 3, 5, 7, 9)
```

I can measure its length with 

```{r subset-1-2, exercise.setup = "subset-1-1", exercise=TRUE}
length(my_vector)
```

Conversely, a data frame will have two dimensions, which I capture with

```{r subset-1-3, exercise=TRUE}
dim(gapminder)
```

The the length of the gapminder data frame is `r ncol(gapminder)` while its "height" is `r nrow(gapminder)` (but "height" is not a function in R, instead use `nrow()` for number of rows).

As seen to subset `my_vector` and extract the first and the third element I can do (replace the `...`)

```{r subset-1-4, exercise.setup = "subset-1-1", exercise = TRUE}
my_vector[...]
```

```{r subset-1-4-solution}
my_vector[c(1, 3)]
```

```{r subset-1-4-check}
grade_this_code()
```

To subset a data frame, I can do something similar but I will need to act on two dimensions, rows and columns, where rows are selected first and columns later, after a comma. For example, to select the 10th row and columns 1, 2 and 3 I can

```{r echo = TRUE}
# Note the comma separating the two conditions for rows and columns respectively
gapminder[10 , c(1,2,3)] 
```

or alternatively if I know the column names

```{r echo = TRUE}
gapminder[10,c("country","continent","year")]
```

But let's say I am interested in subsetting the data frame for data about Afghanistan only disregarding rows about other countries. How can I do it if I don't know the number of rows where records correspond to the country `Afghanistan`? I can express this condition with a logical test before the comma (setting no condition after the comma so to capture all the columns). How would you do it? (Replace the `...` with the correct logical operator)

```{r subset-1-5, exercise = TRUE}
gapminder[gapminder$country ... "Afghanistan", ] 
```

```{r subset-1-5-solution}
gapminder[gapminder$country == "Afghanistan", ] 
```

```{r subset-1-5-check}
grade_this_code()
```


## One continuous and one discrete with geom_boxplot()

Box plots are effective in visualising the distribution of continuous variables across different categories. For example, how life expectancy (`lifeExp`) compares across continents (`continent`) in the `year` 2007?

```{r subset-1-6, exercise = TRUE}
ggplot(data = gapminder[... == ... , ]) + geom_boxplot(aes(x = ..., y = lifeExp)) + labs(title = "Life expectancy in 2007")
```

```{r subset-1-6-solution}
ggplot(data = gapminder[gapminder$year == 2007. , ]) + geom_boxplot(aes(x = continent, y = lifeExp)) + labs(title = "Life expectancy in 2007")
```

```{r subset-1-6-check}
grade_this_code()
```

Note: We can add a title or change the other labels of the chart with `labs(title = "My title", x = "My x-axis label", "y = "My y-axis label")`.

### How do I read a box plot?

<iframe width="560" height="315" src="https://www.youtube.com/embed/zc4ymS5xBQA?si=aXgkMoW_JIXCC7Gu" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture; web-share" referrerpolicy="strict-origin-when-cross-origin" allowfullscreen></iframe>

## One continuous and one discrete with geom_col()

Now let's say you want to visualise GDP per capita by Oceanian countries in 1952. You have two conditions for your data frame. In R, We set these two conditions like this: `condition_1 & condition_2` using `&` between the two conditions. Using these two variables we can say `gapminder$year ... 1952` and `gapminder$continent ... "Oceania"`.

Replace the `...` in the code below to visualise this with `geom_col()` - make sure to add the comma when subsetting the dataset!


```{r subset-1-7, exercise = TRUE}
ggplot(gapminder[...]) + geom_col(aes(x = country, y = gdpPercap))
```

```{r subset-1-7-solution}
ggplot(gapminder[gapminder$year == 1952 & gapminder$continent == "Oceania",]) + geom_col(aes(x = country, y = gdpPercap))
```

```{r subset-1-7-check}
grade_this_code()
```

## Difference between geom_histogram(), geom_bar() and geom_col()

These three functions will produce very similar geometries. Yet, they should be use with different types of variables because they do different things. 

Let's run this code:

```{r geom-1, exercise = TRUE}

ggplot(mpg) + geom_histogram(aes(x = hwy))

```

### `geom_histogram()`

`geom_histogram()` takes only a single variable and this variable must be continuous. By default, `geom_histogram()` will divide your variable into 30 "bins" of equal width (or brackets) and count on one of the two axis how many observations you have in each bin. You can change the number of bins adding `bins = 10` outside of `aes()` but within `geom_histogram()`.

```{r geom-2, exercise = TRUE}

ggplot(mpg) + geom_histogram(aes(x = hwy), bins = 10)

```

Can you see the difference?

### `geom_bar()`

`geom_bar()` also counts how many observations you have in each "bins" but this time bins are defined by the categories (or levels) of a single discrete variable. 

```{r geom-3, exercise = TRUE}

ggplot(mpg) + geom_bar(aes(x = manufacturer))

```

You might have noticed that the bars this time are separated by some space while with `geom_histogram()` they are stacked side-by-side (this to gives the idea that the variable is discrete or continuous). 

### `geom_col()`

`geom_col()` is the function you need to use if you want the height of your columns representing some continuous value already present in your data. In order for the function to work, you need to provide two variables (one discrete, one continuous), for the vertical and horizontal axis. Contrary to `geom_bar()` and `geom_histogram()`, `geom_col()` will not summarise any variable and will return as many columns as observations in the data frame.

```{r geom-4, exercise = TRUE}

ggplot(mpg) + geom_col(aes(x = model, y = cyl))

```

## Continuous function variable

The most common case when you have a continuous function variable is when you have the same observation taken regularly over time.  

```{r echo = TRUE}
gapminder[ , c("year", "country", "lifeExp")]
```

We should consider the variable `lifeExp` as a continuous function of time and the most common way to visualise it is of course with a continuous line (`geom_line()`), running from left (before/past) to right (after/future).

How would you complete this code so to map the evolution of life expectancy (`lifeExp`) over time in Australia?

```{r geom-line-1, exercise = TRUE}
... + geom_line(aes(...))
```

```{r geom-line-1-solution}
ggplot(gapminder[gapminder$country == "Australia",]) + geom_line(aes(x = year, y = lifeExp))
```

```{r geom-line-1-check}
grade_this_code(incorrect = "Not quite right! Try using your R coach on Canvas!")
```

## Scales

The ggplot2 functions starting with `scale_` allows to customise the behaviour of the different scales defined in the aesthetics of the geometries. 

For example,

```{r scale-1, exercise = TRUE}
ggplot(data = gapminder) + 
  geom_point(mapping = aes(x = lifeExp, y = gdpPercap)) +
  scale_y_continuous()
```
 
will produce the same chart as before, indeed `+ scale_y_continuous()` is added by default so including this line in your code will not change the resulting visualisation (try deleting that line (and the `+` at the end of the previous line and run it again). 

Yet I can limit the y-scale to a specific interval adding the attribute `limits = c( ... , ...)`. So to limit the visualisation to observations where the value of y (in this case `gdpPercap`) is between 300 and 1000 dollars (the unit of my variable) I can do

```{r scale-2, exercise = TRUE}
ggplot(data = gapminder) + 
  geom_point(mapping = aes(x = lifeExp, y = gdpPercap)) +
  scale_y_continuous(limits = c(300, 1000))
```

Note that R will throw a warning message saying that with these limits some rows (i.e., observations) have been removed - since they are outside the limits you set. 

How would you limit the scale of the x-axis to observation between 45 and 65 leaving the other y-scale unlimited?

```{r scale-3, exercise = TRUE}

```

```{r scale-3-solution}
ggplot(data = gapminder) + 
  geom_point(mapping = aes(x = lifeExp, y = gdpPercap)) +
  scale_x_continuous(limits = c(45, 65))
```

```{r scale-3-check}
grade_this_code(incorrect = "Not quite right! Try using your R coach on Canvas!")
```

### Scale transformation

A very common problem when visualising the data is when your distribution is highly concentrated around some value with a limited number of outliers. A simple approach to improve the visualisation is to log-transform the problematic scale with `+ scale_x_log10()` or `+ scale_y_log10()`.

Which scale would you transform in this case? (You can run the code to check out the result before submitting) 

```{r  scale-4, exercise = TRUE}
ggplot(data = gapminder) + 
  geom_point(aes(x = lifeExp, y = gdpPercap))
```

```{r scale-4-solution}
ggplot(data = gapminder) + 
  geom_point(aes(x = lifeExp, y = gdpPercap)) + 
  scale_y_log10()
```

```{r scale-4-check}
grade_this_code()
```

## Labels

If you want to modify or add labels to your plot, you can use the `labs()` function. In the example below we are replacing the not-so-descriptive labels on the two axes with something more descriptive. 

```{r  labs-1, exercise = TRUE}
ggplot(data = gapminder) + 
  geom_point(aes(x = lifeExp, y = gdpPercap)) + 
  labs(x = "Life expectancy", y = "GDP per capita")
```


Now try to add the `title` "Gross domestic product and life expectancy" and as `caption` "Data source: Gapminder" (remember, you should always indicate the source of your data!). 

```{r labs-2, exercise = TRUE}
ggplot(data = gapminder) + 
  geom_point(aes(x = lifeExp, y = gdpPercap))  + 
  labs(x = "Life expectancy", y = "GDP per capita",
       ...)
```

```{r labs-2-solution}
ggplot(data = gapminder) + 
  geom_point(aes(x = lifeExp, y = gdpPercap))  + 
  labs(x = "Life expectancy", y = "GDP per capita",
       title = "Gross domestic product and life expectancy", caption = "Data source: Gapminder")
```

```{r labs-2-check}
grade_this_code()
```

## Facets

Facets is a feature offered by ggplot2 to visualise subplots each one representing some specific categorical variables found in your data. For example, compare the plot from this code ...

```{r facets-1, exercise = TRUE}
ggplot(data = gapminder, aes(x = lifeExp)) +
  geom_histogram()
```

with the plot from this code as we added the line `facet_wrap("continent")`.

```{r facets-2, exercise = TRUE}
ggplot(data = gapminder, aes(x = lifeExp)) +
  geom_histogram() +
  facet_wrap("continent")
```

What do you think happened? 

## Submit Report

```{r context="setup"}
submission_ui
```

```{r context="server"}
submission_server()
```
