---
title: "SSPS [4102|6006] Week 01: Introduction and overview"
tutorial:
  id: "week-01"
output:
  learnr::tutorial:
    progressive: true
    ace_theme: github
    theme: united
runtime: shiny_prerendered
description: "This tutorial will introduce you to the basics of R."
---

```{r setup, include=FALSE}
library(gradethis)
library(learnr)
library(sspslearnr)
tutorial_options(exercise.checker = gradethis::grade_learnr, 
                 exercise.reveal_solution="TRUE")
knitr::opts_chunk$set(echo = FALSE)
num_tutorial <- 35
tut_reptitle <- "SSPS [4102|6006] Week 01: In-class Individual Tutorial"
```

## Basics of R

### Running code in RStudio

```{r question-code}
question("How do we run code that has already been written on an R script?",
         answer("We highlight the code we want to run and either (a) manually click on the *Run* icon RStudio displays in the upper right corner of the R script window or (b) use the shortcut *command+enter* in Mac or *ctrl+enter* in Windows", correct = TRUE),
         answer("We highlight the code we want to run and hit *enter*", message = "Almost! You are missing the key that must be used in conjunction with *enter*."),
         answer("We highlight the code we want to run and hit the *spacebar*"),
         answer("We highlight the code we want to run and hit *F1*"),
         post_message = "We highlight the code we want to run and either (a) manually click on the *Run* icon RStudio displays in the upper right corner of the R script window or (b) use the shortcut *command+enter* in Mac or *ctrl+enter* in Windows.",
         random_answer_order = TRUE,
         allow_retry = TRUE
)
```


### R as a calculator


First, we'll learn how to use R as a calculator.

- Use the `+` sign to add 5 and 3

```{r add, exercise=TRUE}

```

```{r add-check}
grade_result(
  pass_if(~ identical(.result, 8))
)
```


- Use the `-` sign to subtract 3 from 5

```{r subtract, exercise=TRUE, exercise.lines = 5}

```

```{r subtract-check}
grade_result(
  pass_if(~ identical(.result, 2))
)
```

- Use the `/` to  divide 6 by 2

```{r divide, exercise=TRUE}

```

```{r divide-check}
grade_result(
  pass_if(~ identical(.result, 6/2))
)
```

- Use the `sqrt()` function to take the square root of 16


```{r sqrt, exercise=TRUE}

```

```{r sqrt-check}
grade_result(
  pass_if(~ identical(.result, sqrt(16)), "Now you know how to use R as a calculator.")
)
```


### Storing results

You can save anything in R to an object. This is handy if you want to reuse some calculation later in your session.

- Replace `...` with the difference `8-1` to save it an object called `mydiff`.
- Print the value of `mydiff` to the console by typing it on its own line.

```{r storing, exercise = TRUE, exercise.lines = 6}
## assign the difference here
mydiff <- ...

## print the value of mydiff on the next line

```

```{r storing-check}
grade_result_strict(
  pass_if(~ identical(mydiff, 7)),
  pass_if(~ identical(.result, 7))
)
```


### Characters and strings


A lot of the time we'll work with numbers in R, but we will also want to use a lot of text. This text can be helpful in producing labels for plots or for labeling categorical variables.

- First, save the text `"social science"` to the variable `course` (replace the first `...`).
- Next, save the text `"learning R"` to the same variable `course` (replace the second `...`).
- Finally, print the value of `course` to the console. What do you think it will say?

```{r strings, exercise=TRUE}
## save the first string
course <- ...

## overwrite the course variable with the second phrase
course <- ...

## print the value of course on the next line

```

```{r strings-check}
grade_result(
  pass_if(~ identical(.result, "learning R"))
)
```

### Copying and reassigning variables

When we assign an existing object to a new name we always make a copy of it. This can be useful when you want it, but it also means you can lose what's in your object if you accidentally overwrite it. Here, we are going to learn about creating a copy of an object before overwriting it.

- Assign the value of `result` to `result2`
- Overwrite the value of `result` with `10 - 2`

```{r copying-setup}
result <- 8 - 2
```

```{r copying, exercise=TRUE}
## First result
result <- 8 - 2

## Assign the value of "result" to "result2"


## Overwrite "result"
result <- ...

## Print result on next line

```

```{r copying-solution}
result2 <- result
```

```{r copying-check}
grade_result_strict(
  pass_if(~ identical(result2, 8 - 2)),
  pass_if(~ identical(.result, 10 - 2)),
  pass_if(~ identical(result, 10 - 2))
)
```

## Working with Data


### Working with real data

Next, we are going to start working with real data: estimates of world population (in thousands). A vector of data called `world.pop` has been loaded with this lesson. The first element is for the year 1950 up to the last element for 2010. You can see that we create the vector by using the `c()` function which concatenates multiple values together into one vector. We enter the data one value at a time, each separated by a comma.

- Print the `world.pop` data by simply typing it into a line of code.

```{r printdata, exercise=TRUE}
## create the world.pop data
world.pop <- c(2525779, 3026003, 3691173, 4449049, 5320817, 6127700, 6916183)

## print the world.pop data

```

```{r printdata-solution}
world.pop
```

### Reading data into your R sessions

```{r question-read}
question("What would be the code to read a CSV file called *testscores.csv* and store the data from that file in an object called *grades*?",
    answer("`grades <- read.csv(''testscores.csv'')`", correct = TRUE),
    answer("`read.csv(''testscores.csv'')`", message = "This code will read the data from the CSV file, but it won't store it into an object called *grades*."),
    answer("`grades <- read.csv(testscores.csv)`", message = "You are almost there. Hint: The file name is text and not (1) the name of an object, (2) the name of a function, (3) the name of an argument, or (4) a special character such as TRUE, FALSE, NA, and NULL."),
    answer("`grades <- read.csv(''testscores'')`", message = "You are almost there. Hint: The file name needs to include its extension (i.e., ''.csv'')."),
    answer("`testcores <- read.csv(''grades.csv'')`", message = "You are almost there. Hint: The name of the file is *testcores.csv* and the name of the object is *grades*, not the other way around."),
    post_message = "The correct answer is: `grades <- read.csv(''testscores.csv'') `<br><br>
    To create an object, we use the assignment operator ` <- ` <br>
    To the left of the assignment operator, we specify the name of the object; *grades* in this case. <br>
    To the right of the assignment operator, we specify the contents, which, in this case, are produced by reading the CSV file *testcores.csv*. <br><br>
    Note: We do not use quotes around the name of an object such as *grades* or around the name of a function such as *read.csv()*, but we do use quotes around the name of a file, because it is text and neither of the exceptions (i.e., it is not (1) the name of an object, (2) the name of a function, (3) the name of an argument, or (4) a special character such as TRUE, FALSE, NA, and NULL).
<br><br> For more information on how to create objects in R, review subsection 1.6.2 of DSS. And, for more information on how to use the `read.csv()` function, review step #2 of the section 1.7 of DSS. In addition, you may want to review the CHEATSHEETs on page 26.",
    random_answer_order = TRUE,
    allow_retry = TRUE
  )
```



### Indexing and subsetting


Vectors are just a series of objects in R that are all stored together in a specific order. What if we want to access a specific value in the vector? Well, for that, we can use the indexing and subsetting tools in R. Specifically, we will use the the square brackets, `[ ]` to access specific values within the vector.


- Use the brackets to access the fourth entry in the `world.pop` data.

```{r worlddata}
world.pop <- c(2525779, 3026003, 3691173, 4449049, 5320817, 6127700, 6916183)
```

```{r indexing1, exercise=TRUE, exercise.setup = "worlddata"}
## access and print the 4th value of world.pop

```

```{r indexing1-check}
grade_result(
  pass_if(~ identical(.result, world.pop[4]))
)
```

- Use the brackets to access and print the first and fourth entries of the `world.pop` data. You will need to use the `c()` function to create a vector of indices that you want to access.


```{r indexing2, exercise=TRUE, exercise.setup = "worlddata"}
## access and print the 1st and 4th value of world.pop

```

```{r indexing2-check}
grade_result(
  pass_if(~ identical(.result, world.pop[c(1,4)]))
)
```

### Using functions

One way we will use R a ton is through functions. Functions are the bread and butter of R. They allow us to act on or get information about vectors and other objects. For instance, the following functions are pretty useful for any vector:

- `length(x)` calculates the number of elements/entries in the `x` vector.
- `min(x)` calculates the smallest value in the `x` vector.
- `max(x)` calculates the largest value in the `x` vector .
- `mean(x)` calculates the average value in the `x` vector (that is the sum of the entries divided by the number of entries).
- `sum(x)` calculates the sum of the values in the `x` vector.

#### Instructions

- Calculate the number of elements in  the `world.pop` vector.

```{r length, exercise = TRUE, exercise.setup = "worlddata"}
## calculate the number of entries in world.pop

```

```{r length-solution}
length(world.pop)
```

```{r length-check}
grade_code()
```

- Calculate  the minimum value of the vector.

```{r min, exercise = TRUE, exercise.setup = "worlddata"}
## calculate the minimum value of world.pop

```

```{r min-solution}
min(world.pop)
```

```{r min-check}
grade_code()
```

- Calculate the average value of the vector.

```{r mean, exercise = TRUE, exercise.setup = "worlddata"}
## calculate the average value of world.pop

```

```{r mean-solution}
mean(world.pop)
```

```{r mean-check}
grade_code()
```

### Replacing values in a vector

Indexing and subsetting allow you to access specific values in the vector, but you can also use the same syntax to replace certain values in the vector. That is we can assign a value such as `x[4] <- 50`, which would replace the fourth entry in the `x` vector with the number 50.

For example, suppose that your research assistant came running in to tell you that the earliest world population data was actually from 1945, not 1950. Here, you will fix this in your vector.

#### Instructions

- Replace the first entry of the `year` vector with 1945.

```{r replace, exercise = TRUE}
# seq() creates a sequence of values
year <- seq(from=1950, to=2010, by = 10)

## update the first entry of year


## print the modified year vector

```

```{r replace-check}
grade_result(
  pass_if(~ identical(year[1], 1945))
)
```


### Arithmetic with vectors

What if we wanted our data in millions of people? How would we create this vector from the vector that we have? (Recall that world.pop is currently in units of thousands of people.) One way would be to do this manually---create a new vector using `c()` that concatenates the world population in millions of people rather than thousands of people. But this is cumbersome, can't we use the vector we already have? Yes!

We can apply many types of arithmetic operators such as addition, subtraction, multiplication, and division to our vector. For example, the code `x + 5` will add the number 5 to each value in the vector. In this exercise, we will create a new vector that is the world population in millions of people, which is just the total population in thousands divided by 1000.

#### Exercise

- Divide the `world.pop` vector by 1000 and assign it to a new vector called `pop.million`.
- Print out the `pop.million` variable.

```{r vector_math, exercise=TRUE, exercise.setup = "worlddata"}
## create the pop.million variable.
pop.million <- ...

## print out the pop.million variable

```

```{r vector_math-check}
grade_result(
  pass_if(~ identical(pop.million, world.pop / 1000))
)
```

### Working with a data.frame

A `data.frame` is an object in R that is basically like a spreadsheet with some number of rows (units) and some number of columns (variables) and a name for each column. There are a number of ways to interact with a `data.frame` to get useful information about it. For example, if I have a `data.frame` called `mydata`, I can do the following:

- `names(mydata)` - returns the column (variable) names of the data.
- `ncol(mydata)` - returns the number of columns in the data.
- `nrow(mydata)` - returns the number of rows of the data.
- `dim(mydata)` - returns a vector of the number of rows and the number of columns (the dimension of the data).
- `summary(mydata)` - provides a summary of each variable in the data.

These are super useful functions Let's use some of these on a data frame, `UNpop`, which has the same information as the `world.pop` vector, but stored as a data frame.

#### Exercise

```{r unpop}
UNpop <- data.frame(
  year = seq(1950, 2010, by = 10),
  world.pop = c(2525779, 3026003, 3691173, 4449049, 5320817, 6127700, 6916183)
)
```

- Print the `UNpop` data frame.

```{r printdf, exercise = TRUE, exercise.setup = "unpop"}
## Print the UNpop data frame

```

```{r printdf-solution}
## Print the UNpop data frame
UNpop
```

```{r printdf-check}
grade_code()
```


- Output the variables names in the `UNpop` data frame.

```{r namesdf, exercise = TRUE, exercise.setup = "unpop"}
## Print the variable names of UNpop

```

```{r namesdf-solution}
names(UNpop)
```

```{r namesdf-check}
grade_code()
```


- Use the `dim` function report the number of rows and columns of the data frame.

```{r dim, exercise = TRUE, exercise.setup = "unpop"}
## Print the dimensions of UNpop

```

```{r dim-solution}
## Print the UNpop data frame
dim(UNpop)
```

```{r dim-check}
grade_code()
```


- Use the `summary` function to show a summary of each variable.

```{r summary, exercise = TRUE, exercise.setup = "unpop"}
## Print a summary of the data in UNpop

```

```{r summary-solution}
summary(UNpop)
```

```{r summary-check}
grade_code()
```


### Subsetting a data frame (I)


You'll often need to access different parts of a data frame to use in other commands. For instance, maybe you want to take the mean of a column of the data frame or maybe you want to see all of the data for the 4th unit. Either way, we'll need to know how to subset the data frame. To select a particular variable from the data frame, you can use the `$` operator. So `mydata$myvar` will be a vector of just the `myvar` column of the `mydata` data frame.


#### Exercise

- Use the `$` to print out the `world.pop` variable from the `UNpop` data frame.

```{r dollar, exercise = TRUE, exercise.setup = "unpop"}
## print out the world.pop variable using $

```

```{r dollar-solution}
UNpop$world.pop
```

```{r dollar-check}
grade_code()
```


- Using the `$` operator, calculate the mean of the `world.pop` variable.

```{r meandollar, exercise = TRUE, exercise.setup = "unpop"}
## calculate the mean world population over this time period

```

```{r meandollar-solution}
mean(UNpop$world.pop)
```

```{r meandollar-check}
grade_code()
```

### Subsetting a data frame (II)

You can also use the brackets `[ ]` to subset from the data frame. But how will R know if you want to subset the rows or the columns? With a data frame as opposed to a vector, you will use a comma and the bracket will have the following form: `[rows, columns]` where the expression before the comma will select the rows and the expression after the comma will select the columns.

- `mydata[,"myvar"]` will select the `myvar` column from `mydata`
- `mydata[1,]` will select the first row of `mydata`
- `mydata[c(1,2,3),]` will select the first three rows of `mydata`
- `mydata[1:3, "myvar"]` will select the first three values of the `myvar` variable of `mydata`

#### Exercise


- Use the bracket commands to extract and print the `world.pop` variable from the `UNpop` data frame.

```{r brackets, exercise = TRUE, exercise.setup = "unpop"}
## use brackets to print out the world.pop variable

```

```{r brackets-solution}
UNpop[, "world.pop"]
```

```{r brackets-check}
grade_code()
```



- Use the bracket commands to extract and print rows 5 through 7 of the `UNpop` data frame.


```{r rows, exercise = TRUE, exercise.setup = "unpop"}
## extract rows 5 through 7 and all variables

```

```{r rows-solution}
UNpop[5:7, ]
```

```{r rows-check}
grade_code()
```


- Use the bracket commands to extract and print rows 5 through 7 of the `world.pop` variable of the `UNpop` data frame.


```{r rowcols, exercise = TRUE, exercise.setup = "unpop"}
## extract values 5 through 7 of the world.pop variable

```

```{r rowcols-solution}
UNpop[5:7, "world.pop"]
```

```{r rowcols-check}
grade_code()
```

## Introductory stats

```{r ch1_topic3_mc}
quiz(caption="",
  question("If *X*  = {10, 4, 6, 8, 22}, what is $\\overline{X}$ ?",
    answer("10", correct = TRUE),
    answer("6", message = "Hint: $\\overline{X}$ stands for the mean of *X* in mathematical notation. And, to compute the mean of a variable, we sum the values across all observations and divide the result by the total number of observations."),
    answer("8", message = "Hint: $\\overline{X}$ stands for the mean of *X* in mathematical notation. And, to compute the mean of a variable, we sum the values across all observations and divide the result by the total number of observations."),
    answer("It is impossible to know with the information given", message = "Hint: $\\overline{X}$ stands for the mean of *X* in mathematical notation. And, to compute the mean of a variable, we sum the values across all observations and divide the result by the total number of observations."),
    post_message = "$\\overline{X}$ stands for the mean of *X* and, here, $\\overline{X}$ equals 10 because: (10+4+6+8+22)/5 = 50/5 = 10.",
    random_answer_order = TRUE,
    allow_retry = TRUE
  ),
  question("If the unit of observation is students and the variable *attended* = {0, 0, 1, 1, 1, 0}, what proportion of students attended class?",
    answer("50 percent of the students attended class", correct = TRUE),
    answer("60 percent of the students attended class", message = "Hint: The proportion of observations that meet a criterion is calculated as: *number of observations that meet criterion* / *total number of observations*. And, to interpret the resulting decimal as a percentage, we multiply it by 100."),
    answer("3 percent of the students attended class", message = "Hint: The proportion of observations that meet a criterion is calculated as: *number of observations that meet criterion* / *total number of observations*. And, to interpret the resulting decimal as a percentage, we multiply it by 100."),
    answer("30 percent of the students attended class", message = "Hint: The proportion of observations that meet a criterion is calculated as: *number of observations that meet criterion* / *total number of observations*. And, to interpret the resulting decimal as a percentage, we multiply it by 100."),
    post_message = "The variable *attended* appears to be binary, where 1 means that the student attended class, and 0 means that the student did not attend class. <br><br> 
    There are six observations in this variable: three 1s and three 0s. This means that three out of the six students attended class, which is a proportion of 50 percent (3/6 = 0.5 and 0.5*100 = 50%).",
    random_answer_order = TRUE,
    allow_retry = TRUE
  ),
question("If the unit of observation is students and *attended* = {0, 0, 1, 1, 1, 0}, what is $\\overline{attended}$ ?",
    answer("$\\overline{attended}$ = 0.5, which we can interpret as `50 percent of the students attended class'", correct = TRUE),
    answer("$\\overline{attended}$ = 0.5, which we can interpret as `students attended 0.5 classes on average'", message = "Hint: Is *attended* a binary or a non-binary variable?"),
    answer("$\\overline{attended}$ = 3, which we can interpret as `students attended 3 classes on average'", message = "Hint: Check your calculations. Plus, is *attended* a binary or a non-binary variable?"),
    answer("$\\overline{attended}$ = 0.3, which we can interpret as `30 percent of the students attended class'", message = "Hint: Check your calculations"),
    post_message = "The correct answer is:
    $\\overline{attended}$ = 0.5, which we can interpret as ''50 percent of the students attended class'' <br><br>
  Calculation: $\\overline{attended}$ ${ }$ equals 0.5 because: (0+0+1+1+1+0)/6 = 3/6 = 0.5. <br><br>
  Interpretation: Here, *attendance* is binary (because it can take only two values: 1s and 0s), so the mean of *attendance* should be interpreted as a proportion and in percentages after multiplying the output by 100 (0.5*100=50%). In particular, it should be interpreted as the proportion of the observations that have the characteristic identified by the variable. In this case, the proportion of students who attended class.<br><br>
  Important: Notice that we arrived to the same conclusion in the prior question. Computing the proportion of 1s in a binary variable is equivalent to computing the mean of the binary variable. Both calculations will lead us to the same conclusion.",
    random_answer_order = TRUE,
    allow_retry = TRUE
  )
)
```

Below is the description of the variables in a dataset from a survey of students participating to Project Star, where the unit of observation is students.</b>

<table>
  <tr> 
    <th> variable </th>
    <th> description </th>
  </tr>
  <tr>
    <td> *classtype* </td>
    <td>  class size the student attended: 'small' or 'regular'
  </tr>
  <tr>  
    <td>*reading* </td>
    <td> student's third-grade reading test scores (in points) </td>
  </tr>
  <tr>  
    <td> *math* </td>
    <td>  student's third-grade math test scores (in points) </td>
  <tr>
    <td>*graduated* </td>
    <td>  identifies whether the student graduated from high school: <br>1=graduated or 0=did not graduate </td>
  </tr>
</table>
 
<br> 
 
**SETUP: Suppose we have already set the working directory to the DSS folder and we have already run the code below to read and store the STAR dataset in an object called `star`.** 

```{r star_FALSE2, eval=FALSE, echo=TRUE}
star <- read.csv("STAR.csv") # reads and stores data
```

```{r star, echo=FALSE}
star <- read.csv("data/STAR.csv")
```

**Go ahead and run the following piece of code to ask R to show the first six observations in the dataframe stored in the `star` object:**
 
```{r ch1_topic3_r1_setup, exercise=TRUE, exercise.setup = "star"}
head(star) # shows the first six observations
```

**Use the `mean()` function to compute the mean of the variable *reading***

```{r ch1_topic3_r1, exercise=TRUE, exercise.setup = "star"}

```

<div id="ch1_topic3_r1-hint">
Hint: *reading* is a variable of the dataframe stored inside the object `star`, and the `$` character might be helpful when accessing variables of dataframes.
</div>

```{r ch1_topic3_r1-solution}
mean(star$reading)
```

```{r ch1_topic3_r1-check}
grade_code()
```

```{r ch1_topic3_r1_inter}
quiz(caption="",
       question("How can we interpret the mean of *reading* computed above?",
    answer("Among the students who participated in Project STAR, the average score in the third-grade reading test is about 629 points", correct = TRUE),
    answer("Among the students who participated in Project STAR, the average score in the third-grade reading test is about 629%", message="Check the unit of measurement of *reading*"),
    answer("Among the students who participated in Project STAR, the average score in the third-grade reading test is about 629", message="You are missing the unit of measurement."),
    answer("Among the students who participated in Project STAR, the average score in the third-grade reading test is about 628 points", message="Notice that we round up whenever the next decimal point is 5 or above."),
    post_message = "The correct answer is: Among the students who participated in Project STAR, the average score in the third-grade reading test is about 629 points.<br><br>
Here, as we can see in the output of `head()` above, *reading* is non-binary because it can take more than two values. And, as we can read in the description of the variable in the table above, *reading* is measured in points. So, the mean of *reading* should be interpreted as an average and in points.",
    random_answer_order = TRUE,
    allow_retry = TRUE
  ))
```

**Now, use the `mean()` function to compute the mean of the variable *graduated* **

```{r ch1_topic3_r2, exercise=TRUE, exercise.setup = "star"}

```

<div id="ch1_topic3_r2-hint">
Hint: *graduated* is a variable of the dataframe stored inside the object `star`, and the `$` character might be helpful when accessing variables of dataframes.
</div>

```{r ch1_topic3_r2-solution}
mean(star$graduated)
```

```{r ch1_topic3_r2-check}
grade_code()
```

```{r ch1_topic3_r2_inter}
quiz(caption="",
       question("How can we interpret the mean of *graduated* computed above?",
    answer("87% of the students who participated in Project STAR graduated from high school", correct = TRUE),
    answer("0.87% of the students who participated in Project STAR graduated from high school", message="Remember that to interpret the output as a percentage, we need to multiply it by 100. (Review how to interpret the mean of binary variable in section 1.8 of DSS.)"),
    answer("87 students in Project STAR graduated from high school.", message="Is *graduated* a binary or a non-binary variable?"),
    answer("Fewer than one student in Project STAR graduated from high school", message="Is *graduated* a binary or non-binary variable?"),
    post_message = "The correct answer is: 87% of the students who participated in Project STAR graduated from high school.<br><br>
   Here, as we can see in the output of `head()` above and read in the description of variables in the table above, *graduated* is binary because it can take only two values: 1s and 0s. So, the mean of *graduated* should be interpreted as a proportion and in percentages after multiplying the output by 100 (0.87*100=87%). In particular, it should be interpreted as the proportion of the observations that have the characteristic identified by the variable. In this case, the proportion of students who graduated from high school.",
    random_answer_order = TRUE,
    allow_retry = TRUE
  ))
```

<br> <span style=color:dodgerblue> **FIX THE ERRORS IN THE CODE:** </span> **Running into error messages is part of coding, so let's practice fixing mistakes in our code. There are many mistakes in the code below. Try to fix them all so that R computes de mean of the variable *math*, which is one of the variables in the dataframe stored in the object called *star*. (The error messages that R provides can be really helpful when looking to debug code. So, I recommend starting by running the code as is, and then running it again after fixing each individual error, so that you can read the messages that R provides.)**

```{r ch1_topic3_r3, exercise=TRUE, exercise.setup = "star"}
average(math$STAR) # computes the mean of math
```

<div id="ch1_topic3_r3-hint">
Hints: <br><br>
The function that calculates the mean of a variable is `mean()`, not `average()`.<br><br>
To the left of the `$` character, we specify the name of the object where the dataframe is stored (without quotes). To the right of the `$` character, we specify the name of the variable (without quotes). Example: `data$variable`. <br><br>
R code is case sensitive and the name of the object where we have stored the dataset is `star`, not `STAR`.
</div>

```{r ch1_topic3_r3-solution}
mean(star$math)
```

```{r ch1_topic3_r3-check}
grade_code("<br><br>The function that calculates the mean of a variable is `mean()`, not `average()`.<br><br>
We use the `$` character to access a variable inside a dataframe. To its left, we specify the name of the object where the dataframe is stored (without quotes). To its right, we specify the name of the variable (without quotes). Example: `data$variable`. <br><br>
R code is case sensitive and the name of the object where we have stored the dataset is `star`, not `STAR`.")
```

```{r ch1_topic3_r3_inter}
quiz(caption="",
       question("How can we interpret the mean of *math* computed above?",
    answer("Among the students who participated in Project STAR, the average score in the third-grade math test is about 632 points", correct = TRUE),
    answer("Among the students who participated in Project STAR, the average score in the third-grade math test is about 63.2%", message="Check the unit of measurement"),
    answer("Among the students who participated in Project STAR, the average score in the third-grade math test is about 632", message="You are missing the unit of measurement."),
    answer("Among the students who participated in Project STAR, the average score in the third-grade math test is about 631 points", message="Notice that we round up whenever the next decimal point is 5 or above."),
    post_message = "The correct answer is: Among the students who participated in Project STAR, the average score in the third-grade math test is about 632 points.<br><br>
     Here, as we can see in the output of `head()` above, *math* is non-binary because it can take more than two values. And, as we can read in the description of the variable in the table above, *math* is measured in points. So, the mean of *math* should be interpreted as an average and in points.",
    random_answer_order = TRUE,
    allow_retry = TRUE
  ))
```

## Submit Report

```{r context="setup"}
submission_ui
```

```{r context="server"}
submission_server()
```
